---
title: "311 Service Calls of New York City"
author: "Yisong Tao"
date: "10/14/2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

311 is a public service phone number of many communities in North America, which provides access to many non-emergency municipal services[[1]][Wikipedia]. New York City now supports both phone call and online interfaces to 311 service, and provide open access to 311 service requests data from 2010 to present. There are several well-written articles online utilizing this dataset to come up with insights into data-driven urban management[[2]][Wired] or specific issues in NYC[[3]][NewYorker]. 

The questions I want to answer by analyze and visualization of New York 311 data are: what are the top complaints? Are they the same across 5 boroughs? Try to characterize the top complaints.

```{r set working directory load libraries, warning=FALSE, message=FALSE}
setwd("~/Desktop/Project1")
library(dplyr)
library(ggplot2)
library(ggthemes)
```

The NYC 311 service requests data can be accessed from  [https://nycopendata.socrata.com/Social-Services/311-Service-Requests-from-2010-to-Present/erm2-nwe9/data][Link]. The dataset for this project was downloaded on 10/01/2016, which was a 9.15 Gb csv file containing over 137 million requests from 2010-01-01 to 2016-09-30 02:11:54. 

```{r download data, eval=FALSE}
fileurl <- "https://nycopendata.socrata.com/api/views/erm2-nwe9/rows.csv?accessType=DOWNLOAD"
file <- "data.csv"
download.file(fileurl, file, method = "curl" )
data <- data.table::fread(file, sep = ",", header = TRUE, 
                          stringsAsFactors = FALSE, 
                          na.strings = c("N/A", "", "NA", "Unspecified"))
nrow(data)           #dataset contains 13720953 observations
names(data)          #dataset contanins 53 variables
saveRDS(data, "all_data.RDS")
data_311 <- readRDS("all_data.RDS")
```

## 1. What are the most complained issues in New York?
```{r fig. 1 Complained Types, cached= T, echo=T, dev='png', fig.path='figures/', fig.width= 10, fig.height= 8}
#data_311 <- readRDS("all_data.RDS")
#all_complaints <- data_311 %>%select(6, 25) #Select "Complaint Type" and "Borough" columns
#saveRDS(all_complaints, "all_complaints.RDS")            
all_complaints <- readRDS("all_complaints.RDS")
all_complaints$`Complaint Type`[grepl("^Noise.*", all_complaints$`Complaint Type`)] <- "Noise" 
all_complaints <- all_complaints %>%
    group_by(Borough, `Complaint Type`) %>% summarise(Count = n())
all_complaints_NY <- all_complaints %>%
    group_by(`Complaint Type`) %>%
    summarise(Count = sum(Count)) %>%
    arrange(desc(Count))
top10_complaints_NY <- top_n(all_complaints_NY, 10, Count)
figure1 <- function(df){
    return(ggplot(df) + geom_bar(aes(x=reorder(`Complaint Type`,Count) , y = Count),
             stat = "identity") + theme_few() +
    xlab("") + ylab("Number of Complaints") + coord_flip() )
}
p_top10_complaints_NY  <- figure1(top10_complaints_NY) + ggtitle("All 5 Boroughs")
top10_complaints_borough <- top_n(group_by(all_complaints, Borough), 10, Count)
top10_complaints_borough <- arrange(top10_complaints_borough, desc(Borough))
top10_complaints_Man <- filter(top10_complaints_borough, Borough == "MANHATTAN")
p_top10_complaints_Man  <- figure1(top10_complaints_Man) + ggtitle("Manhattan")
top10_complaints_Qns <- filter(top10_complaints_borough, Borough == "QUEENS")
p_top10_complaints_Qns  <- figure1(top10_complaints_Qns) + ggtitle("Queens")
top10_complaints_Bn <- filter(top10_complaints_borough, Borough == "BROOKLYN")
p_top10_complaints_Bn  <- figure1(top10_complaints_Bn) + ggtitle("Brooklyn")
top10_complaints_Brx <- filter(top10_complaints_borough, Borough == "BRONX")
p_top10_complaints_Brx  <- figure1(top10_complaints_Brx) + ggtitle("Bronx")
top10_complaints_SI <- filter(top10_complaints_borough, Borough == "STATEN ISLAND")
p_top10_complaints_SI  <- figure1(top10_complaints_SI) + ggtitle("Staten Island")
#### The following function allos multiple graphs in one plot, the following multiplot 
#### function is from http://www.cookbook-r.com/Graphs/Multiple_graphs_on_one_page_(ggplot2)/
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)
  plots <- c(list(...), plotlist)
  numPlots = length(plots)
  if (is.null(layout)) {
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }
  if (numPlots==1) {
    print(plots[[1]])
  } else {
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    for (i in 1:numPlots) {
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
####
multiplot(p_top10_complaints_NY, p_top10_complaints_Man, p_top10_complaints_Bn,
          p_top10_complaints_Qns, p_top10_complaints_Brx, p_top10_complaints_SI,
          cols=2)
```
```{r percentage function, echo=F}
percent <- function(x, digits = 2, format = "f", ...) {
  paste0(formatC(100 * x, format = format, digits = digits, ...), "%")
}
```


The above barplot shows the top 10 complained issues from New York City and each of the 5 boroughs. Noise is the most complained issue in New York City(`r percent(all_complaints_NY$Count[1]/sum(all_complaints_NY$Count))` of all complaints) and each borough except for Staten Island, where noise is the second most complained after street condition. Heating, street light condition and street condition are also among the most complained issue.

##2.  Regarding noise complaints, which neighborhood/zip codes are making most of  the noises in New York City?

```{r noise complaints analysis, eval=F}
noise_complaints <- select(data_311, 6, 7, 9, 25, 51, 52)
noise_complaints$Descriptor <- as.factor(noise_complaints$Descriptor)
noise_complaints <- filter(noise_complaints, grepl("Noise", `Complaint Type`))
noise_complaints$Type <- NA
noise_complaints$Type[grepl("Dog|Animals",
                            noise_complaints$Descriptor)] <- "Dog and other animals"
noise_complaints$Type[grepl("Air Condition|air condition",
                            noise_complaints$Descriptor)] <- "Air Conditioner"
noise_complaints$Type[grepl("Banging", 
                            noise_complaints$Descriptor)] <- "Banging/Pounding"
noise_complaints$Type[grepl("Truck|Vehicle|Boat|Private Carting|Engine|Flying|Hovering|Honking", noise_complaints$Descriptor)] <- "Vehicle"
noise_complaints$Type[grepl("Construction|Jack Hammering|Manufacturing", noise_complaints$Descriptor)] <- "Construction"
noise_complaints$Type[grepl("Loud Music|Television|Talking", 
                            noise_complaints$Descriptor)] <- "Music/TV/Talking"
noise_complaints$Type[grepl("Alarm", 
                            noise_complaints$Descriptor)] <- "Alarm"
noise_complaints$Type[is.na(noise_complaints$Type)] <- "Other"
noise_complaints<-select(noise_complaints, 2, 3, 4, 5, 6, 7)
noise_complaints$`Incident Zip` <- substr(noise_complaints$`Incident Zip`, 1, 5)
nyc_list <- readRDS("nyc_list.RDS") ## load a list of NYC zip codes
noise_complaints <- filter(noise_complaints, `Incident Zip` %in% nyc_list)
saveRDS(noise_complaints, "noise_complaints.RDS")

```

```{r noise complaints analysis2}
noise_complaints <- readRDS("noise_complaints.RDS")
noise_sum<- as.data.frame(table(noise_complaints$Descriptor))
noise_sum <- noise_sum[noise_sum$Freq > 0,]
noise_sum <- noise_sum[order(-noise_sum$Freq),]
```
There are 36 different descriptors for all `r nrow(noise_complaints) ` noise complaints.
Top 10 complained noise types are:
```{r noise type, echo=F}
head(noise_sum, 10)
```
I grouped the 36 types into 8 types to simplify the plots. These 8 types of noises are _Construction_, _Dog and other animals_, _Vehicle_, _Music/TV/Talking_, _Alarm_, _Air Conditioner_, _Banging/Pounding_ and _Other_. The following analysis also uses the population data from 2010 census to normalize the results. Due to the low population in certain zip codes, some of the per-capita values had to removed as out-liers.

```{r fig. 2_1 Noise Complaints spatial distribution, cached= T, echo=T, dev='png', fig.path='figures/', fig.width= 10, fig.height= 8, warning=F, message=F}
## load NY state population data by zip code from 2010 census
nyc_pop <- read.csv("aff_download/DEC_10_SF1_P1_with_ann.csv", skip = 1,
                    header = T, stringsAsFactors = F) 

nyc_pop$Zip <- as.factor(nyc_pop$Id2)
noise_complaints <- readRDS("noise_complaints.RDS")
library(leaflet)
library(tmap)
nyczipgeo <- readRDS("nyczipgeo.RDS") ## load NYC zip code shape map
noise_sum_zipcode <- as.data.frame(table(noise_complaints$`Incident Zip`))
noise_sum_zipcode$Zip <- noise_sum_zipcode$Var1
noise_sum_zipcode <- left_join(noise_sum_zipcode, nyc_pop)
noise_sum_zipcode <- noise_sum_zipcode %>% select(Zip, Freq, Total) %>%
    mutate( Count = Freq/Total)
noise_sum_zipcode$Count[noise_sum_zipcode$Count == Inf] <- NA
noise_sum_zipcode$Count[which.max(noise_sum_zipcode$Count)] <- NA
nycmap <- append_data(nyczipgeo, noise_sum_zipcode, key.shp = "ZCTA5CE10", key.data = "Zip")
nyc_map<- tm_shape(nycmap) +
    tm_fill("Count", title = "All Noise", palette = "YlOrRd") +
    tm_borders(alpha = 0.5) +
    tm_style_natural(legend.frame = F, legend.bg.color = NA)

noise_sum_zipcode_type <- noise_complaints %>%
    group_by(`Incident Zip`, Type) %>%
    summarise(Count = n())
noise_sum_zipcode_type <- reshape2::dcast(noise_sum_zipcode_type, `Incident Zip` ~ Type)
noise_sum_zipcode_type <- left_join(noise_sum_zipcode_type, nyc_pop,
                                    by =c("Incident Zip" = "Zip"))
noise_sum_zipcode_type$`Air Conditioner` <- noise_sum_zipcode_type$`Air Conditioner`/ noise_sum_zipcode_type$Total
noise_sum_zipcode_type$`Air Conditioner`[noise_sum_zipcode_type$`Air Conditioner`== Inf] <- NA
noise_sum_zipcode_type$`Air Conditioner`[which.max(noise_sum_zipcode_type$`Air Conditioner`)] <- NA
noise_sum_zipcode_type$Alarm <- noise_sum_zipcode_type$Alarm/ noise_sum_zipcode_type$Total
noise_sum_zipcode_type$Alarm[noise_sum_zipcode_type$Alarm== Inf] <- NA
noise_sum_zipcode_type$Vehicle <- noise_sum_zipcode_type$Vehicle/ noise_sum_zipcode_type$Total
noise_sum_zipcode_type$Vehicle[noise_sum_zipcode_type$Vehicle== Inf] <- NA
noise_sum_zipcode_type$Construction <- noise_sum_zipcode_type$Construction/ noise_sum_zipcode_type$Total
noise_sum_zipcode_type$Construction[noise_sum_zipcode_type$Construction== Inf] <- NA
noise_sum_zipcode_type$Construction[which.max(noise_sum_zipcode_type$Construction)] <- NA
noise_sum_zipcode_type$`Banging/Pounding` <- noise_sum_zipcode_type$`Banging/Pounding`/ noise_sum_zipcode_type$Total
noise_sum_zipcode_type$`Banging/Pounding`[noise_sum_zipcode_type$`Banging/Pounding`== Inf] <- NA
noise_sum_zipcode_type$`Dog and other animals` <- noise_sum_zipcode_type$`Dog and other animals`/ noise_sum_zipcode_type$Total
noise_sum_zipcode_type$`Dog and other animals`[noise_sum_zipcode_type$`Dog and other animals`== Inf] <- NA
noise_sum_zipcode_type$`Music/TV/Talking` <- noise_sum_zipcode_type$`Music/TV/Talking`/ noise_sum_zipcode_type$Total
noise_sum_zipcode_type$`Music/TV/Talking`[noise_sum_zipcode_type$`Music/TV/Talking`== Inf] <- NA
noise_sum_zipcode_type$Other <- noise_sum_zipcode_type$Other/ noise_sum_zipcode_type$Total
noise_sum_zipcode_type$Other[noise_sum_zipcode_type$Other== Inf] <- NA

nyc_noise_map <- append_data(nyczipgeo, noise_sum_zipcode_type,
                             key.shp = "ZCTA5CE10", key.data = "Incident Zip") 
nyc_noise_map_construction <- tm_shape(nyc_noise_map) + 
    tm_fill("Construction", title = "Construction", palette = "YlOrRd") +
    tm_borders(alpha = 0.5) + 
    tm_style_natural(legend.frame = F, legend.bg.color = NA)
nyc_noise_map_dogs <- tm_shape(nyc_noise_map) +
    tm_fill("Dog and other animals",
            title = "Dog and other Animals", palette = "YlOrRd") +
    tm_borders(alpha = 0.5) +
    tm_style_natural(legend.frame = F, legend.bg.color = NA)
nyc_noise_map_vehicle <- tm_shape(nyc_noise_map) +
    tm_fill("Vehicle", title = "Vehicle", palette = "YlOrRd") +
    tm_borders(alpha = 0.5) +
    tm_style_natural(legend.frame = F, legend.bg.color = NA)
nyc_noise_map_music <- tm_shape(nyc_noise_map) +
    tm_fill("Music/TV/Talking", title = "Music/TV/Talking",
            palette = "YlOrRd") +
    tm_borders(alpha = 0.5) +
    tm_style_natural(legend.frame = F, legend.bg.color = NA)
nyc_noise_map_alarm <- tm_shape(nyc_noise_map) +
    tm_fill("Alarm", title = "Alarm", palette = "YlOrRd") +
    tm_borders(alpha = 0.5) +
    tm_style_natural(legend.frame = F, legend.bg.color = NA)
nyc_noise_map_ac <- tm_shape(nyc_noise_map) +
    tm_fill("Air Conditioner", title = "Air Conditioner",
            palette = "YlOrRd") + tm_borders(alpha = 0.5) +
    tm_style_natural(legend.frame = F, legend.bg.color = NA)
nyc_noise_map_other <- tm_shape(nyc_noise_map) +
    tm_fill("Other", title = "Other", palette = "YlOrRd") +
    tm_borders(alpha = 0.5) +
    tm_style_natural(legend.frame = F, legend.bg.color = NA)

nyc_noise_map_banging <- tm_shape(nyc_noise_map) +
    tm_fill("Banging/Pounding",
            title = "Banging/Pounding", palette = "YlOrRd") +
    tm_borders(alpha = 0.5) +
    tm_style_natural(legend.frame = F, legend.bg.color = NA)
nyc_map
```
The neighborhoods complaining about noises most are northern Manhattan, mid Manhattan and lower Manhattan, an area in Brooklyn(zip codes 11205, 11216, 11217 and 11238) and several neighborhoods in Bronx.
```{r fig. 2_2, cached= T, echo=T, dev='png', fig.path='figures/', fig.width= 10, warning=F, message=F}
multiplot(nyc_noise_map_music, nyc_noise_map_banging, cols=2)
```
"Loud Music/Party" is the most complained type of noise in NYC, accounting for over 50% of all noise complaints. It's not surprising that its spatial distribution is very similar to that of over all noise. "Banging/Pounding" is the second most complained type. It is most severe in the neighborhood of mid Manhattan(Zip code 10001) and Upper Manhattan (zip codes 10030 an 10039).  
```{r fig. 2_3, cached= T, echo=T, dev='png', fig.path='figures/', fig.width= 10, warning=F, message=F}
multiplot(nyc_noise_map_construction, nyc_noise_map_vehicle, cols=2)
```
Complaints about construction noises are most common in mid and low Manhattan, while vehcile noises are most complained around mid, low  and northern Manhattan.
```{r fig. 2_4, cached= T, echo=T, dev='png', fig.path='figures/', fig.width= 10, warning=F, message=F}
multiplot(nyc_noise_map_ac, nyc_noise_map_other, cols=2)
```
Air condition and ventilation noises are most complained in mid and low Manhattan. The category "Other" is mostly complained near JFK air port, Red Hook neighborhood of Brooklyn and Washington Heights of Manhattan.
```{r fig. 2_5, cached= T, echo=T, dev='png', fig.path='figures/', fig.width= 10, warning=F, message=F}
multiplot(nyc_noise_map_dogs, nyc_noise_map_alarm, cols=2)
```
Complaints about barking dogs and other animals are relatively rare in NYC and it spreads across all boroughs. It is mostly common in the neighborhood(zip code 10464) near Westchester County. Complaints about "Alarm" are very peculiar as the neighborhood around Kew Gardens (zip code 11366) contributed near 20% of the complaints, alomst 10x more than any other zip codes in NYC.

##3. How about other common complaints such as "Heating" and "Street Condition"?
```{r fig. 3_1 Heating and Street Condition,cached= T, echo=T, dev='png', fig.path='figures/', fig.width= 10, warning=F, message=F}
#other_complaints <- all_complaints%>%select(2, 6, 7, 9, 25, 51, 52)
#other_complaints$`Created Date` <- as.POSIXct(strptime(data_311$`Created Date`, "%m/%d/%Y %I:%M:%S %p"))
#other_complaints$Month <- format(other_complaints$`Created Date`, "%b")
#other_complaints$`Incident Zip` <- substring(other_complaints$`Incident Zip`, 1, 5)
#saveRDS(other_complaints, "other_complaints.RDS")
#other_complaints <- readRDS("other_complaints.RDS")
#Heating <- filter(other_complaints, grepl("HEATING", other_complaints$`Complaint Type`))
#Heating$Month <- as.factor(Heating$Month)
#Heating <- select(Heating, -1)
#saveRDS(Heating, "Heating.RDS")
Heating <- readRDS("Heating.RDS")
Heating_plot <- ggplot(data = Heating) + 
    theme_few() + scale_fill_few() + ylab("Number of Complaints") +
    geom_bar(aes(x= Month, fill = Borough), stat = "count") + theme(legend.position = "top") +
    scale_x_discrete(limits = c ("Jan", "Feb", "Mar", "Apr", "May","Jun",
                                 "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

#StreetCondition <- filter(other_complaints, grepl("Street Condition", other_complaints$`Complaint Type`))
#StreetCondition$Month <- as.factor(StreetCondition$Month)
#StreetCondition <- select(StreetCondition, -1)
#saveRDS(StreetCondition, "StreetCondition.RDS")
StreetCondition <- readRDS("StreetCondition.RDS")
StreetCondition_plot <- ggplot(data = StreetCondition) + 
    theme_few() + scale_fill_few() + ylab("Number of Complaints") +
    geom_bar(aes(x= Month, fill = Borough), stat = "count") +
    scale_x_discrete(limits = c ("Jan", "Feb", "Mar", "Apr", "May","Jun",
                                 "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) +
    theme(legend.position = "none")

multiplot(Heating_plot, StreetCondition_plot, cols = 2)
```
Complaints about heating is most often in cold weather form October to April. This is as expected.
The situation is similar with complaints about street conditions, which is most complained from Februry to Jun, after all the long and cold NY winter time.
```{r fig. 3_2 Heating and Street Condition,cached= T, echo=T, dev='png', fig.path='figures/', fig.width= 10, warning=F, message=F}
nyc_list <- readRDS("nyc_list.RDS") ## load a list of NYC zip codes
Heating_sum <- Heating %>% filter(`Incident Zip` %in% nyc_list) %>%
    group_by(`Incident Zip`) %>% summarise(Count = n())
heating_map <- append_data(nyczipgeo, Heating_sum,
                           key.shp = "ZCTA5CE10", key.data = "Incident Zip")
heating_map_plot <- tm_shape(heating_map) +
    tm_fill("Count", title = "Heating Complaints", palette = "YlOrRd") +
    tm_borders(alpha = 0.5) +
    tm_style_natural(legend.frame = F, legend.bg.color = NA)
StreetCondition_sum <- StreetCondition %>%
    filter(`Incident Zip` %in% nyc_list) %>%
    group_by(`Incident Zip`) %>% summarise(Count = n())
StreetCondition_map <- append_data(nyczipgeo, StreetCondition_sum,
                                   key.shp = "ZCTA5CE10", key.data = "Incident Zip")
StreetCondition_map_plot <- tm_shape(StreetCondition_map) +
    tm_fill("Count", title = "Street Condition Complaints",
            palette = "YlOrRd") + tm_borders(alpha = 0.5) +
    tm_style_natural(legend.frame = F, legend.bg.color = NA)

multiplot(heating_map_plot, StreetCondition_map_plot, cols = 2)
```
It appears some neighborhoods in Brooklyn and Bronx suffers most heating problems. Interestingly, these neighborhoods coincide with the "Banging/Pounding" noise complaints map. Maybe the deteriorating heating infrastrcutre in these neighborhoods has some connection with the "banging" noise. 
As for road conditions, Staten Island is hit most hard, many neighborhoods in Brooklyn and Queens also suffer a lot. Manhattan and Bronx seem to have less prolem with the bad street conditions.


## Summary:
After exploratory analysis of NYC 311 dataset, we conclude that New Yorkers' most complained issue is noise, most of which are from Manhattan and some of its adjacent neighborhoods in Bronx and Brooklyn. Heating and street condition are also often complained, these two complaints have their own correlation to the cold weather. Their time and spatial pattern could be useful for the city to distribute the resources to help resolving these issues.

## References:
[1. https://en.wikipedia.org/wiki/3-1-1][Wikipedia]  
[2. What a Hundred Million Calls to 311 Reveal About New York - Wired][Wired]  
[3. Mapping New York’s Noisiest Neighborhoods - The New Yorker][NewYorker]  

[Wikipedia]:https://en.wikipedia.org/wiki/3-1-1  
[Wired]:https://www.wired.com/2010/11/ff_311_new_york/  
[NewYorker]:http://www.newyorker.com/tech/elements/mapping-new-york-noise-complaints  
[Link]:https://nycopendata.socrata.com/Social-Services/311-Service-Requests-from-2010-to-Present/erm2-nwe9/data
