---
title: "Music Explorer"
author: "Joshua Litven"
date: "10/30/2016"
output: ioslides_presentation
---

## Question

*How do you learn about new music?*

## Question

*How do you learn about new music?*
![title](https://static1.squarespace.com/static/547dce0ce4b037bc72d6db93/t/55d23c1de4b0a6ed8765a26b/1439841310475/?format=750w)

## Motivation

- Exploring different genres of music
- Learning about artists
- Filtering large playlists

## Shiny App: Music Explorer

Let's see it in action!

## Challenges

- Scraping data
- Constructing the artist network graph
- Audio using Javascript

## Scraping Data

```{r eval=FALSE}
get_playlists = function(category_id){
  # Request playlists
  URI = paste0('https://api.spotify.com/v1/browse/categories/', 
               category_id, '/playlists')
  request = GET(url = URI, add_headers(Authorization = HeaderValue))
  playlist_object = fromJSON(content(request, "text")) 
  playlists = playlist_object$playlists$items
  return(playlists)
  
}
get_tracks = function(playlist_id, user_id){
  # Request track features
  URI = paste0('https://api.spotify.com/v1/users/',
               user_id,'/playlists/', playlist_id,'/tracks')
  response2 = GET(url = URI, add_headers(Authorization = HeaderValue))
  tracks = fromJSON(content(response2, "text"))$items$track
  return(tracks)
}
```

## Constructing the Artist Network

- For each artist in the playlist, create a node
- Take the union of artists and their related artists
- Define an edge to connect related artists
- Filter out artists with only one edge
- Result: All artists are connected to at least two other artists

## Audio in Graphs using Javascript

```{r eval=FALSE}
histogram_js = "if(chart.getSelection().length > 0){
                    var row_index = chart.getSelection()[0].row;
                    var tooltip = data.getValue(row_index, 0);
                    Shiny.onInputChange('tooltip',tooltip);
                  }else{Shiny.onInputChange('tooltip', '');}"
gvisHistogram(data.frame(name=tracks$full_name, minutes=minutes),
                  options=list(title="Track Duration",
                               titleTextStyle="{fontName:'Courier', 
                               fontSize:24}",
                               legend="none",
                               hAxis="{title: 'Minutes', 
                                      titleTextStyle:{fontSize:20}}",
                               tooltip="{isHtml:'True'}",
                               gvis.listener.jscode = histogram_js))
```

## Insights

- Networks consists of subnetworks
    + Connected by crossover artists
- "*Lone wolf*" nodes
    + Artists that do not connect to other artists
- Artist tags are strange
    + e.g. "*contemporary post-bop*" "*post-teen pop*", "*dirty south rap*", "*crunk*"
- Some Spotify metrics seem dubious
    + e.g. beats per minute, mood, liveness
- Others are more consistent
    + e.g. energy, decibals, popularity

## Future Work

- Add more playlists
- Functionality to input any public playlist
- Spotify login to view private playlists
- Cluster network to identify subgenres
- Improve performance

Your turn!!
https://jlitven.shinyapps.io/music_explorer/


