---
title: "What's Hot?"
subtitle: "Analysis of Top 10 Reddit Posts"
author: "Regan Yee"
date: "November 12, 2016"
output: ioslides_presentation
css: assets/css/ioslides.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(plotly)
library(dplyr)
library(DT)
library(png)
library(grid)
setwd("~/Documents/bootcamp007_project/Project3-WebScraping/ReganYee/Wayback")
reddit = readRDS("reddit.RDS")

df = reddit %>% 
  mutate(datehour = sapply(reddit$snapshot_datetime, function(x) substring(x,1,10))) %>% 
  mutate(rank_hour = paste0(datehour,rank)) %>% 
  select(1:10, rank_hour, submit_date, submit_hour)

## Return a list of top 10 items that are unique per day and hour
df_unique_hour = df[!duplicated(df$rank_hour),]
df_unique_hour = df_unique_hour %>% select(snapshot_date, rank, titles, subreddit, upvotes, comments, submitter, snapshot_time, url, submit_datetime, submit_date, submit_hour)

## Reindex row indices
rownames(df_unique_hour) = NULL

df_unique_hour = df_unique_hour %>%
  mutate(snapshot_hour = substring(snapshot_time,1,2))
```

## Questions
<b>Scraping:</b>

- How can I scrape historical data for time-series data?

<b>Data:</b>

- What subreddits produce the most front-page content?
- What are people talking about on Reddit?
- When is the best time to submit content?

## Data 
<a href="https://archive.org/web/"> The Wayback Machine </a> is a site that allows us to browse historical snapshots of any websites crawled by the site.

I <a href="https://raw.githubusercontent.com/nycdatasci/bootcamp007_project/master/Project3-WebScraping/ReganYee/Wayback/wayback.csv">scraped</a> ~211k rows of data about the top 10 Reddit posts from 06/30/2016 - 11/11/2016:

```{r reddit_snippet, echo=TRUE}
colnames(reddit)
```

<div>
<font size="-1">Upvotes had data randomly imputed whenever there was an upvote value of '•'.</font>
<br/>
<font size="-1">Comments for threads which restricted comments had a value of 0 imputed.</font>
</div>

## Scraping Workflow
```{r workflow}
img = readPNG("./assets/img/snap1.png")
grid.raster(img)
```

## {.flexbox .vcenter}
<div class="centered">
<font size="+5">Findings</font>
</div>

## Hottest Subreddits by Posts
```{r subreddit_analysis, cache = FALSE, echo = FALSE, message=FALSE}
subreddit_grouped = reddit %>% group_by(subreddit,titles) %>%
  summarize(count=n()) %>% arrange(desc(count))

## Filter on subreddits with more than 10 posts in top10
subreddit_analysis = subreddit_grouped %>% 
  select(subreddit) %>% 
  summarize(count=n()) %>% 
  filter(count >=10) %>% 
  arrange(desc(count))

## Reorder to be descending
subreddit_analysis$subreddit = 
  factor(subreddit_analysis$subreddit, levels = unique(subreddit_analysis$subreddit)[order(subreddit_analysis$count, decreasing = TRUE)])

y <- list(
  title = "Number of Posts")

plot_ly (
  x = subreddit_analysis$subreddit,
  y = subreddit_analysis$count,
  type = 'bar') %>% 
  layout(
    yaxis = y
  )
```
<div>
<font size="-1">Only including subreddits with >= 10 posts in the top 10</font>
</div>

## Comments vs Upvotes
```{r reddit_comment_ratio, cache = FALSE, echo = FALSE, message=FALSE}
comment_analysis1 = reddit %>% 
  group_by(Titles=titles,Subreddit=subreddit) %>% 
  summarize(Ratio = max(signif(comment_ratio, 5))) %>% 
  arrange(desc(Ratio))

datatable(head(comment_analysis1, 8),
          extensions = 'FixedColumns',
          options = list(
            dom = 't',
            scrollX = TRUE,
            scrollCollapse = TRUE)) %>% 
  formatStyle('Titles', `font-size` = '14px') %>% 
  formatStyle('Subreddit', `font-size` = '14px') %>% 
  formatStyle('Ratio', `font-size` = '14px')
```

## Hottest Upvote Time (by Date)
```{r reddit_score_plot, cache = FALSE, echo = FALSE, message=FALSE}
reddit_score_mean = df_unique_hour %>% 
  select(snapshot_date,upvotes) %>% 
  group_by(snapshot_date) %>% 
  summarize(avg_score_mean = mean(upvotes))

model = lm(avg_score_mean ~ as.numeric(snapshot_date), data = reddit_score_mean)

plot_ly (
  x = reddit_score_mean$snapshot_date,
  y = reddit_score_mean$avg_score_mean,
  type = 'scatter',
  mode = 'markers',
  xaxis = 'x1') %>% 
  add_trace(x = as.numeric(reddit_score_mean$snapshot_date), y = fitted(model), mode='lines', xaxis = 'x2') %>%
  layout(
    xaxis=list(
        xaxis = 'x1',
        title = "Date",
        nticks = 4
    ),
    xaxis2=list(
        overlaying='x1',
        showticklabels = FALSE),
    yaxis = list(
      title = "Reddit Score"
    )
  )
```

## Hottest Upvote Time (by Snapshot Hr)
```{r reddit_score_hour, cache = FALSE, echo = FALSE, message=FALSE}
reddit_score_by_hour = df_unique_hour %>% 
  group_by(snapshot_hour) %>% summarize(score_mean = mean(upvotes), score_median = median(upvotes))

x <- list(
  title = "Hour of day(UTC)")
y <- list(
  title = "Reddit Score")

plot_ly (
  x = reddit_score_by_hour$snapshot_hour,
  y = reddit_score_by_hour$score_mean,
  type = 'scatter',
  mode = 'lines',
  name = 'mean') %>% 
  add_trace(
    x = reddit_score_by_hour$snapshot_hour,
    y = reddit_score_by_hour$score_median,
    type = 'scatter',
    mode = 'lines',
    name = 'median'
  ) %>% 
  layout(
    xaxis = x,
    yaxis = y
  )
```

## Hottest Upvote Time (by Submit Time)
```{r reddit_score_submitted, cache = FALSE, echo = FALSE, message=FALSE}
reddit_score_by_submit_hour = df_unique_hour %>% 
  group_by(submit_hour) %>% summarize(score_mean = mean(upvotes), score_median = median(upvotes))

x <- list(
  title = "Hour of day(UTC)")
y <- list(
  title = "Reddit Score")

plot_ly(
    x = reddit_score_by_submit_hour$submit_hour,
    y = reddit_score_by_submit_hour$score_mean,
    type = 'scatter',
    mode = 'lines',
    name = 'mean') %>% 
  add_trace(
    x = reddit_score_by_submit_hour$submit_hour,
    y = reddit_score_by_submit_hour$score_median,
    type = 'scatter',
    mode = 'lines',
    name = 'median') %>% 
  layout(
    xaxis = x,
    yaxis = y
  )
```

## Conclusions
- Reddit users gravitate towards r/funny, r/pics, r/gifs, r/todayilearned, r/aww, and r/gaming.
- r/AskReddit is the subreddit that generates the most discussion via the comments.
- There appears to be a positive trend in upvotes as the days go by in my dataset.
- The best time to submit content to Reddit to get the most amount of coverage is 13:00 - 19:00 UTC.
- The posts on reddit are most likely to have the most upvotes from 17:00 - 20:00 UTC.

## Retrospective
- Scraping Wayback Machine site is only good for sites which have enough snapshots per day.
- Website layouts change and hence need to be able to handle different tag name changes in the spider.
- Data became missing at random due to HTTP 429. Would like to find a way to try again for all the items that reach a 429 status.

##
<div class="centered">
<font size="+5">Questions?</font>
</div>

