---
title: "Vehicular Collisions in NYC"
author: "Regan Yee"
date: "October 12, 2016"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
setwd("~/Documents/Project/")
vehc = read.csv("NYPD_Motor_Vehicle_Collisions.csv")
library(RgoogleMaps)
library(dplyr)
library(ggplot2)
library(ggmap)
library(choroplethrZip)

############### By Time
# We want to organize the crashes by hours, so let's strip hours from TIME field
date_hour = strptime(vehc$TIME,"%H")
hour = as.numeric(format(date_hour, "%H")) +
  as.numeric(format(date_hour, "%M"))/60
vehc$hour = hour

# Create a monthDate column
v = strsplit(as.character(vehc$DATE),"/")
v1 = matrix(unlist(v), ncol=3, byrow=TRUE)
v2 = paste(v1[,1],v1[,2],sep="/")
vehc$monthDate = v2

# Create a month column
v = strsplit(as.character(vehc$DATE),"/")
v1 = matrix(unlist(v), ncol=3, byrow=TRUE)
vehc$month = v1[,1]

# Create a year column
v = strsplit(as.character(vehc$DATE),"/")
v1 = matrix(unlist(v), ncol=3, byrow=TRUE)
vehc$year = v1[,3]

# Organize a df sorted by the number of crashes in an hour slot
common_times = vehc %>% group_by(hour) %>%
                        summarize(count=n()) %>% 
                        arrange(desc(count))

# Organize a df sorted by the number of crashes in an month
common_months = vehc %>% group_by(month) %>%
  filter(year %in% c(2013,2014,2015)) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))

############### By Year
## In a list v, delimit the DATE by "/" and strip out the YEAR
v = strsplit(as.character(vehc$DATE),"/")
v1 = matrix(unlist(v), ncol=3, byrow=TRUE)

############### Contributing Factors Summaries
contributing_factors = vehc %>% select(CONTRIBUTING.FACTOR.VEHICLE.1,CONTRIBUTING.FACTOR.VEHICLE.2,CONTRIBUTING.FACTOR.VEHICLE.3,CONTRIBUTING.FACTOR.VEHICLE.4,CONTRIBUTING.FACTOR.VEHICLE.5)
distinct(contributing_factors, CONTRIBUTING.FACTOR.VEHICLE.1)
sum1 = contributing_factors %>% group_by(CONTRIBUTING.FACTOR.VEHICLE.1) %>% summarize(count=n())
sum2 = contributing_factors %>% group_by(CONTRIBUTING.FACTOR.VEHICLE.2) %>% summarize(count=n())
sum1 %>% arrange(desc(count))
sum2 %>% arrange(desc(count))

############### Number of people killed by year sum(Pedestrians, Cyclists, Motorists)
vehc %>% group_by(year) %>% summarize(total=sum(NUMBER.OF.PERSONS.KILLED))
colnames(vehc)
vehc %>% group_by(year) %>% summarize(total=sum(NUMBER.OF.PEDESTRIANS.KILLED))

zip_counts = vehc %>% 
  group_by(year,ZIP.CODE) %>%
  summarize(count = n()) %>%
  select(year, ZIP.CODE, value = count) %>%
  mutate(region = as.character(ZIP.CODE))

crashes_in_2016 = zip_counts %>% filter(year==2016)
crashes_in_2015 = zip_counts %>% filter(year==2015)
crashes_in_2014 = zip_counts %>% filter(year==2014)
crashes_in_2013 = zip_counts %>% filter(year==2013)
crashes_in_2012 = zip_counts %>% filter(year==2012)

# New York City is comprised of 5 counties: Bronx, Kings (Brooklyn), New York (Manhattan), 
# Queens, Richmond (Staten Island). Their numeric FIPS codes are:
nyc_fips = c(36005, 36047, 36061, 36081, 36085)

g2016 = zip_choropleth(crashes_in_2016,
               county_zoom=nyc_fips,
               num_colors = 1)
## 2015 vs 2014
y2015.vs.y2014 = left_join(crashes_in_2015,crashes_in_2014,by = 'region')
df20152014 = as.data.frame(y2015.vs.y2014) %>% select(year.x,region,value.x,value.y)
df20152014 = df20152014 %>% mutate(value=value.x-value.y)

# Choropleth map of changes from 2014-2015
diff2015 = zip_choropleth(df20152014,
                       county_zoom=nyc_fips,
                       num_colors = 1)
## 2014 vs 2013
y2014.vs.y2013 = left_join(crashes_in_2014,crashes_in_2013,by = 'region')
df20142013 = as.data.frame(y2014.vs.y2013) %>% select(year.x,region,value.x,value.y)
df20142013 = df20142013 %>% mutate(value=value.x-value.y)

# Choropleth map of changes from 2015-2014
diff2014 = zip_choropleth(df20142013,
                          county_zoom=nyc_fips,
                          num_colors = 1,
                          legend="Difference in Accidents")

# Fatalities by Borough
people_killed_by_borough = vehc %>% group_by(year,BOROUGH) %>% summarize(total=sum(NUMBER.OF.PERSONS.KILLED))

```

## Dataset Summary
[NYPD_Motor_Vehicle_Collisions](https://data.cityofnewyork.us/Public-Safety/NYPD-Motor-Vehicle-Collisions/h9gi-nx95) provided by NYC Open Data

This dataset provides insights into the all vehicular collisions within the NYC area

- 902679 observations with 30 variables
- Updated on weekdays (Current data is 07/01/2012 - 10/11/2016)
- Data sourced from the NYPD

## What are we trying to answer?

- Is Vision Zero working? i.e. Is there a decreasing trend?
- Where are the highest number of collisions?
- When? (by Month, by Hour of day)
- Common causes for collisions

## YTD 2016 Vehicular Crashes

```{r 2016crashes, cache = FALSE, echo = FALSE, message=FALSE}
g2016 + scale_fill_continuous(
  low = "#FFFFFF",
  high = "red",
  space = "Lab",
  na.value = "grey50",
  guide = "colourbar",
  limits = c(0, 3000),
  name = "Number of Crashes"
  )
```

Generated using choroplethrZip library

## Changes from 2013 to 2014

```{r diff2014, cache = FALSE, echo = FALSE, message=FALSE}
diff2014 + scale_fill_gradient2(
  low = "#67a9cf",
  mid= "#f7f7f7",
  high = "#ef8a62",
  midpoint = 0,
  space = "Lab",
  na.value = "white",
  name = "Diff",
  limits = c(-275,200)
)
```

2013 n = 203689; 2014 n = 205929



## Changes from 2014 to 2015

```{r diff2015, cache = FALSE, echo = FALSE, message=FALSE}
diff2015 + scale_fill_gradient2(
  low = "#67a9cf",
  mid= "#f7f7f7",
  high = "#ef8a62",
  midpoint = 0,
  space = "Lab",
  na.value = "white",
  name = "Diff"
)
```
2014 n = 205929; 2015 n = 217539  


## Number of Fatal Crashes per Year
``` {r fatal, echo = FALSE, eval = FALSE}
vehc %>% group_by(year) %>% summarize(total=sum(NUMBER.OF.PERSONS.KILLED))
```

Year | Count of Fatalities
------------- | -------------
2013 | 297
2014 | 262
2015 | 243

## Fatalities by Borough
``` {r by_borough, echo = FALSE}
ggplot(people_killed_by_borough %>% filter(BOROUGH != "") %>% filter(year %in% c(2013,2014,2015))) + geom_bar(aes(x=year,y=total),stat="identity") + facet_wrap( ~ BOROUGH)
```

## Vehicular Crashes by Hour of Day

```{r hours, cache = TRUE, echo = FALSE, fig.height=3, fig.ext='png'}
g_hour=ggplot(data=common_times)
g_hour+geom_bar(aes(x=hour, y=count),stat="identity") + xlab("Hour of Day") + ylab("Count Crashes")
```

## Vehicular Crashes by Month
```{r months, cache = TRUE, echo = FALSE, fig.height=3, fig.ext='png'}
g_month=ggplot(data=common_months)
g_month+geom_bar(aes(x=month, y=count),stat="identity") + xlab("Month") + ylab("Count Crashes")
```

Data from full years of 2013, 2014, 2015


## Top 5 Contributing Factors
Number | Reason | Count
------------- | ------------- | -------------
1 | Unspecified | 500826
2 | Driver Inattention/Distraction | 106757
3 | Fatigued/Drowsy | 48153
4 | Failure to Yield Right-of-Way | 36776
5 | Other Vehicular | 27953

15 - Alcohol Involvement  
20 - Aggressive Driving/Road Rage  
42 - Cell Phone (hands-free)  
43 - Cell Phone (hand-held)

## Top 5 Contributing Factors - Fatal Collisions
Number | Reason | Count
------------- | ------------- | -------------
1|Unspecified|536
2|Traffic Control Disregarded|124
3|Driver Inattention/Distraction|95
4|Failure to Yield Right-of-Way|75
5|Passenger Distraction|52

## Conclusions
- Vehicular collisions are on the rise, but fatalities have been decreasing since 2013.
- The Vision Zero initiative may be benefiting certain neighborhoods more than others.
- Collisions most commonly occur during the the hours of 4-6PM.
- Collisions happen most frequently around Manhattan's Midtown, East New York/Carnarsie, and Long Island City.

## Next Steps
- Compare NYC's stats to other cities
- Look at other cities who are doing a Vision Zero initiative and benchmark NYC's progress
- Compare neighborhood median incomes and see if there is a correlation to the improvement from Vision Zero initiatives




