---
title: "Data Visualization"
author: "Jordan Waters"
date: "October 16, 2016"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
allbyweek <- read.csv('allbyweek.csv', header = T)
allbyweek <- as.data.frame(allbyweek)
ncaa <- read.csv('ncaa.csv', header = T)
ncaa <- as.data.frame(ncaa)
library(ggplot2)
```

## Better Betting with Time?

- Millions of bets are placed every year on college football games
- As betters gain more insight about teams, do they predict the outcomes more accurately?
- Betting trends can be infered from what casinos call "Money lines".
- Money lines exist to create equal betting on both teams in a game. Without getting into too much detail, casinos pay out more to people who bet on teams that most betters think will lose.

## Week 1 Money lines

```{r allbyweek, echo = TRUE}
ggplot(filter(as.data.frame(allbyweek),allbyweek$Week == 1 & !is.na(allbyweek$Team_Line)), aes(x = reorder(Team, Team_Line), y = Team_Line)) + geom_bar(stat = 'identity', aes(fill = win)) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle("All Games from Week 1") + ylab('Money Line') + xlab('Team')

```

## Slide with Plot

```{r pressure}
plot(pressure)
```

