---
title: "Annual Payroll Study"
author: "Connie Zhang"
date: "October 15, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading the libraries needed for the project
````{r, load library}
library(ggplot2)
library(varhandle)
library(dplyr)
library(dygraphs)
library(googleVis)
library(RColorBrewer)
`````

## Overview of the Project

The goal of the project is to invistigate the distribution of US payroll from 2005 to 2014 in order to explore the envolution of business patterns over this period. 

The questions we try to answer are:

1. How the annual payroll evolved during the period and is there any trend?
2. Does the geographic factor have impact on the payroll distribution?
3. How the payrolls are developed among the different industries?

The data used in the study is published by US Census Bureau and the downloaded from the link:

http://factfinder.census.gov/faces/tableservices/jsf/pages/productview.xhtml?pid=BP_2014_00A1&prodType=table

The data we are interested in contains the following information:

1. State: 50 US states and district of Columbia
2. Industry:  21 classifications of industries defined by NAICS code
3. Employees: this is the sample collected from the first quart payroll during the year.
4. Aggregate payroll: for the each industry defined above

The following data is not completed or available to the industries below:

Agriculture, forestry, fishing and hunting
Mining
Utilities (mostly)
Industries not classified

The project will ignore this missing data and remove it when necessary.

## loading the data into dataframe and combine them in to list
````{r, load data}
bp05 = read.csv("~/data/BP05.csv")
bp06 = read.csv("~/data/BP06.csv")
bp07 = read.csv("~/data/BP07.csv")
bp08 = read.csv("~/data/BP08.csv")
bp09 = read.csv("~/data/BP09.csv")
bp10 = read.csv("~/data/BP10.csv")
bp11 = read.csv("~/data/BP11.csv")
bp12 = read.csv("~/data/BP12.csv")
bp13 = read.csv("~/data/BP13.csv")
bp14 = read.csv("~/data/BP14.csv")

bp_data = list(bp05, bp06,bp07, bp08, bp09, bp10,
               bp11, bp12, bp13, bp14)
````
## Convert factor class into numerical for payroll and employees
````{r, for unfactor}
for ( i in 1:length(bp_data)) {
  bp_data[[i]]$Paid.employees.for.pay.period.including.March.12..number. = 
    unfactor(bp_data[[i]]$Paid.employees.for.pay.period.including.March.12..number.)
  bp_data[[i]]$Annual.payroll...1.000. = unfactor(bp_data[[i]]$Annual.payroll...1.000.)
  class(bp_data[[i]]$Annual.payroll...1.000.)
  class(bp_data[[i]]$Paid.employees.for.pay.period.including.March.12..number.)
}
````
## US Payroll Over time

````{r, for US Payroll over time}
Years = c(rep.int(0, length(bp_data)))
PaidEmployee = c(rep(0,length(bp_data)))
AnnalPay = c(rep(0,length(bp_data)))

for (i in 1:length(bp_data)) {             
Total =filter(bp_data[[i]], NAICS.code == "0") %>% 
  summarise(Total.paid.Employee =sum(Paid.employees.for.pay.period.including.March.12..number.),
            Total.Annual.payroll.thousand.dollar = sum(Annual.payroll...1.000.))
Years[i] = unique(bp_data[[i]]$Year)
PaidEmployee[i] = Total$Total.paid.Employee
AnnalPay[i] = Total$Total.Annual.payroll.thousand.dollar/1.0e+6
}
avgAnnalPay = AnnalPay/PaidEmployee *1.0e+9
AnnPayroll_ts = ts(AnnalPay, start=2005,end=2014, frequency = 1)
avgAnnalPay_ts = ts(avgAnnalPay, start=2005,end=2014, frequency = 1)
PaidEmployee_scaled = PaidEmployee/(1.0e+6)
PaidEmployee_ts = ts(PaidEmployee_scaled,start=2005,end=2014,frequency =1)


dygraph(AnnPayroll_ts, main="Annual Payroll 2005-2014 (Bilions)") %>% 
              dySeries(label = "Payroll", color = "black") %>%
              dyShading(from ="2007-01-01",to = "2010-01-01",color ="#CCEBD6")

dygraph(avgAnnalPay_ts, main="Average Annual Paid 2005-2014") %>% 
  dySeries(label = "Annual Average Payroll", color = "black") %>%
  dyShading(from ="2007-01-01",to = "2010-01-01",color ="#CCEBD6")

dygraph(PaidEmployee_ts, main="Paid Employees (millions) 2005-2014") %>% 
  dySeries(label = "Paid Employees", color = "black") %>%
  dyShading(from ="2007-01-01",to = "2010-01-01",color ="#CCEBD6")
```````
## The Average payroll distribution over states in 2007, 2009 and 2014
`````{r Payroll distribution on states}
infor2007=filter(bp_data[[3]],NAICS.code == "0")
infor2009=filter(bp_data[[5]],NAICS.code == "0")
infor2014=filter(bp_data[[10]],NAICS.code == "0")

infor2007$avgAnnualPayroll = infor2007$Annual.payroll...1.000./infor2007$Paid.employees.for.pay.period.including.March.12..number.*1000
infor2009$avgAnnualPayroll = infor2009$Annual.payroll...1.000./infor2009$Paid.employees.for.pay.period.including.March.12..number.*1000
infor2014$avgAnnualPayroll = infor2014$Annual.payroll...1.000./infor2014$Paid.employees.for.pay.period.including.March.12..number.*1000
````
## Average Annual Payroll distribution in 2007
```{r, results= 'asis'}
plot(gvisGeoChart(infor2007, "Geographic.area.name", "avgAnnualPayroll",
                               options=list(region="US", 
                               displayMode="regions", 
                               resolution="provinces",
                               width=800, height = 600,
                               colorAxis="{colors: ['green', 'white', 'red']}")),"chart"  ) 
````
## Average Annual Payroll Distribution in 2009
```` {r, results= 'asis'}    
plot(gvisGeoChart(infor2009, "Geographic.area.name", "avgAnnualPayroll",
                       options=list(region="US", 
                                    displayMode="regions", 
                                    resolution="provinces",
                                    width=800, height = 600,
                                    colorAxis="{colors: ['green', 'white', 'red']}")), "chart" )   

````
## Average Annual Paryroll Distribution in 2014
`````{r, results= 'asis'}
plot(gvisGeoChart(infor2014, "Geographic.area.name", "avgAnnualPayroll",
                            options=list(region="US", 
                                         displayMode="regions", 
                                         resolution="provinces",
                                         width=800, height = 600,
                                         colorAxis="{colors: ['green', 'white', 'red']}")), "chart"  ) 
`````
## boxplot for the national wide distribution for 2007, 2009 and 2014
`````{r, for boxplot}
Year = c(infor2007$Year,infor2009$Year, infor2014$Year)
AvgAnnalPayRoll = c(infor2007$avgAnnualPayroll,infor2009$avgAnnualPayroll,
                    infor2014$avgAnnualPayroll)
avgPayroll = data.frame(Year, AvgAnnalPayRoll)
avgPayroll$Year = factor(avgPayroll$Year)
avgPayboxplot = ggplot(data = avgPayroll, aes(x = Year, y=AvgAnnalPayRoll))
avgPayboxplot +geom_boxplot() +
  ggtitle("Average AnnualPay in state level:2007,2009,2014")
`````
## Payroll variance among industries within states
````{r adding column and calculate the standard deviation}
for(i in 1:length(bp_data)) {
bp_data[[i]]$avgAnnualPayroll = ifelse((is.na(bp_data[[i]]$Paid.employees.for.pay.period.including.March.12..number.)
                           | is.na(bp_data[[i]]$Annual.payroll...1.000.)),NA,
 bp_data[[i]]$Annual.payroll...1.000./bp_data[[i]]$Paid.employees.for.pay.period.including.March.12..number.*1000)
}

# variation of avgAnnualPayroll among industries within states

PayrollStd = list ()

for (i in 1:length(bp_data)){

PayrollStd[[i]] = bp_data[[i]] %>% group_by(State =Geographic.area.name) %>% 
                summarise(STD =sd(avgAnnualPayroll, na.rm = TRUE))
}
````
## Average Payroll standard deviation in 2007
`````{r, results= 'asis'}
plot(gvisGeoChart(PayrollStd[[3]],"State","STD",
                               options=list(region="US", 
                                            displayMode="regions", 
                                            resolution="provinces",
                                            width=800, height = 600,
                                            colorAxis="{colors: ['green', 'white', 'red']}")), "chart")                  

`````
## Average Payroll standard deviation in 2009
`````{r, results= 'asis'}

plot(gvisGeoChart(PayrollStd[[5]],"State","STD",
                              options=list(region="US", 
                                           displayMode="regions", 
                                           resolution="provinces",
                                           width=800, height = 600,
                                           colorAxis="{colors: ['green', 'white', 'red']}")), "chart")                  


`````
## Average Payroll standard deviation in 2014
`````{r, results= 'asis'}
plot(gvisGeoChart(PayrollStd[[10]],"State","STD",
                              options=list(region="US", 
                                           displayMode="regions", 
                                           resolution="provinces",
                                           width=800, height = 600,
                                           colorAxis="{colors: ['green', 'white', 'red']}")), "chart")                  

````
## Average Payroll Based on Industries
We will look at the average payroll for each industries over years, and
we also compute the top 3 and tail 3 industries each year based on average payroll across states.
````{r average payroll distribution on industries}
by_industry =list ()

for (i in 1:length(bp_data)){

by_industry[[i]] =bp_data[[i]] %>% filter(NAICS.code != 0) %>% group_by(Meaning.of.NAICS.code) %>%
  summarise(TotalAnnualPayroll = sum(Annual.payroll...1.000.),
                       TotalPaidEmployees = sum(Paid.employees.for.pay.period.including.March.12..number.))
by_industry[[i]]$avgAnnualPayroll =ifelse((is.na(by_industry[[i]]$TotalAnnualPayroll) | 
                                        is.na(by_industry[[i]]$TotalPaidEmployees)),NA,
                            by_industry[[i]]$TotalAnnualPayroll*1000/by_industry[[i]]$TotalPaidEmployees)
}
# Merging average payroll info over years

AvgPayOverYear = data.frame(Industry =by_industry[[1]]$Meaning.of.NAICS.code,
                            Yr.2005=by_industry[[1]]$avgAnnualPayroll)
for (i in 2:length(bp_data)) {
  ii = i+1
  AvgPayOverYear[ii] = by_industry[[i]]$avgAnnualPayroll
  
}
colnames(AvgPayOverYear) = c('Industry','Yr.2005','Yr.2006','Yr.2007','Yr.2008',
                          'Yr.2009','Yr.2010','Yr.2011','Yr.2012','Yr.2013','Yr.2014')

## get industries of top 3 and tail 3 paid for this time period
  
tt = list()
tt1 = list()
Top3= list()
Tail3 = list()

for (i in 1:length(bp_data)) {
  j=i+1
  Top3overYears = AvgPayOverYear[j] %>% top_n(3)
  Tail3overYears = AvgPayOverYear[j] %>% top_n(-3)
  Top_vec=Top3overYears[,1]
  Tail_vec = Tail3overYears[,1]
  Vec_AvgPayOverYear = AvgPayOverYear[,j]
  tt[[i]] = filter(AvgPayOverYear, Vec_AvgPayOverYear %in% Top_vec)
  Top3[[i]]=data.frame(Industry=tt[[i]]$Industry,tt[[i]][j])
  tt1[[i]] = filter(AvgPayOverYear, Vec_AvgPayOverYear %in% Tail_vec)
  Tail3[[i]]=data.frame(Industry =tt1[[i]]$Industry,tt1[[i]][j])
}

## plot the top 3 counts over the years
  
Top3Names = as.character(Top3[[1]]$Industry)

for (i in 2:length(bp_data)) {

 temp_char =as.character(Top3[[i]]$Industry)
 Top3Names = c(Top3Names,temp_char)  
}
Top3Names = factor(Top3Names)
Top3_ct = table(Top3Names)
barplot(Top3_ct,names.arg=c("Fin & Ins","Info",
                            "Mgmt of Entpr.",
                            "Prof.Serv","Utilites"))

## Plot the top and tail Average payroll overtime
# Time series for top and tail  over time
FinInsure=filter(AvgPayOverYear,Industry == "Finance and insurance")
FinInsure =FinInsure[,2:11]
FinInsure=as.numeric(FinInsure)
Information = filter(AvgPayOverYear,Industry == "Information")
Information =Information[,2:11]
Information = as.numeric(Information)

Retail=filter(AvgPayOverYear,Industry == "Retail trade")
Retail=Retail[,2:11]
Retail=as.numeric(Retail)
AccomFood = filter(AvgPayOverYear,Industry == "Accommodation and food services")
AccomFood = AccomFood[,2:11]
AccomFood = as.numeric(AccomFood)

Year = seq(from =2005, to=2014,by=1)
Year = factor(Year)
OvertimeTopTail = data.frame(Year, FinInsure,Information,Retail,AccomFood)

ggplot(OvertimeTopTail, aes(x=Year, y = value,color = variable)) +
  geom_point(aes(y = FinInsure, col="FinInsure")) +
  geom_point(aes(y=Information,col ="Information")) +
  geom_point(aes(y=Retail,col="Retail")) +
  geom_point(aes(y=AccomFood,col="Accom & Food")) +
  ggtitle("Average Pay over time for the Top and Tail Industries") +
  ylab("Average Paid Amount") +
  theme_bw() 
 
```````

#   Conclusion

Based on the study above, the following information is shown:

1. US payroll overall has been increasing over years since 2009. We do see the recession had the most serious impact on employment but less impact on average payroll nationalwide.

2. North east and West coast are the regions to have big payroll increase over years compared with central and they also have larger payroll variance accross industries within state.

3. Finance and insurance, information are the top average paid accross United states. Retail, Accomodation and Food are the poorest average paid.The average paid from top performers increase more quickly than that of tail ones to increase the wealth gap over years. And the trend seems to stay in the future.  
 

