---
title: "Explore House Sale Data"
author: "Ziqiao (Cheryl) Liu"
date: "10/16/2016"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(ggplot2)
library(scales)
library(RColorBrewer)
library(ggthemes)

house_0= read.csv("~/Downloads/train.csv")

house_1 = filter(house_0, GrLivArea<4000)

names = c("MSZoning",
          "HouseStyle", "OverallQual", "YearBuilt","Foundation",
          "MoSold", "YrSold", "SalePrice")

house_data = house_1[, names]

n_1 = c('2Story', '1Story', '1.5Fin')
n_2 = c('2.5Unf','2.5Fin','1.5Unf','SFoyer','SLvl')

house_n_1=house_data%>%filter(.,HouseStyle%in%n_1)
house_n_2=house_data%>%filter(.,HouseStyle%in%n_2)

```

## Outlines

- Background of data
- Insights explored from data
- Reference: <https://www.kaggle.com/c/house-prices-advanced-regression-techniques/data>.

## Background of Data

- The data set describes the sale of individual residential property in Ames, Iowa from 2006 to 2010.

- The data selected for this visualization project (8 attritubes, 1456 observations) includes Categorial data, Continuous data.

- "*The variables focus on the quality and quantity of many physical attributes of the property. Most of the variables are exactly the type of information that a typical home buyer would want to know about a potential property.*"[1]

* [1]"Ames, Iowa: Alternative to the Boston Housing Data as an End of Semester Regression Project", Dean De Cock, Truman State University

## Background of Data

- 8 attritubes selected, 1456 observations
- **Sale Price** (in $)
- **Year Built** : original construction date
- **Year Sold** 
- **Month Sold** 
- **House Style**: style of dwelling
- **House Location Zone**: identifies the general zoning classification of the sale. 
- **Foundation**: type of foundation
- **Overall Quality** (from 1 to 10)

## Background of Data

```{r house_data, echo = TRUE}
summary(house_data)
```

## Prices Trend with House Built Year

```{r , echo = F}
g1 = ggplot(data= house_data, aes(x= YearBuilt, y =SalePrice))
plot1 = g1 + geom_point(aes(color=HouseStyle),position = "jitter", pch=8, cex=.8)+
  geom_smooth(method='lm',size = 0.5, se = T) + 
  scale_y_continuous(labels = comma) +
  ggtitle(" House Sale Price Trend ") +
  ylab("Sale Price ($)") +
  xlab("Year Built") +
  scale_color_manual(values = c('#fc8d62', '#dfc27d', '#7fbc41', '#80cdc1', '#a6cee3', '#4eb3d3', '#984ea3', '#e7298a'),
                     name = "House Style", labels=c("1.5-Story", "1.5-Story(U)", "1-Story",
                                                    "2.5-Story", "2.5-Story(U)","2-Story",
                                                    "Split Foyer", "Split Level"))
plot1 

```

## House Prices: Popular House Styles
```{r, echo=F}
g2_1 = ggplot(data= house_n_1, aes(x= YearBuilt, y =SalePrice, color = HouseStyle))
plot2_1 = g2_1 + geom_point(aes(color=HouseStyle),position = "jitter", pch=8, cex=.3)+
  geom_smooth(method='lm',size = 1.0, se = FALSE)+ 
  scale_y_continuous(labels = comma) +
  theme(legend.position ="bottom")+
  ggtitle(" House Sale Price Trend for House Styles") +
  ylab("Sale Price ($)") +
  xlab("House Built Year") +
  scale_color_manual(values = c( '#fc8d62','#7fbc41', '#4eb3d3'),
                     name = "House Style", labels=c("1.5-Story","1-Story",
                                                     "2-Story"))
plot2_1

```


## House Prices: Discontinued ... and Emerged House Styles
```{r}
g2_2 = ggplot(data= house_n_2, aes(x= YearBuilt, y =SalePrice, color = HouseStyle))
plot2_2 = g2_2 + geom_point(aes(color=HouseStyle),position = "jitter", pch=8, cex=.3)+
  geom_smooth(method='lm',size = 1.0, se = FALSE)+ 
  scale_y_continuous(labels = comma) +
  ggtitle(" House Sale Price Trend for House Styles") +
  ylab("Sale Price ($)") +
  xlab("House Built Year") +
  scale_color_manual(values = c('#dfc27d', '#80cdc1', '#a6cee3', '#984ea3', '#e7298a'),
                     name = "House Style", labels=c("1.5-Story(U)", "2.5-Story", "2.5-Story(U)",
                                                    "Split Foyer", "Split Level"))
plot2_2
```

## House Prices Trends with Sold Years
```{r, echo = F}

house_yrsold= house_data%>%select(.,YrSold, SalePrice)%>%
  group_by(.,YrSold)%>%summarise(.,total = n(), median_p=median(SalePrice))
house_yrsold$YrSold <- factor(house_yrsold$YrSold, levels=reorder(house_yrsold$YrSold, -house_yrsold$total))
g3 = ggplot(data=house_yrsold, aes(x = YrSold, y = total)) 
plot3 = g3 + geom_bar(aes(fill=median_p), stat = "identity") + 
  scale_fill_gradient(guide="colorbar",high = "#132B43", low = "#56B1F7", name= "Median Price($)")+
  geom_text(aes(label = median_p ), size = 4,  hjust = 0.5, vjust = -0.3)+
  xlab(" ") +
  ylab("Number of Houses Sold")+
  ggtitle(" House Sale vs. Year")
plot3

```

## Houses Price Trends with Sold Month 
```{r, echo= F}
house_mosold= house_data%>%select(.,YrSold, MoSold,SalePrice)%>%
  group_by(.,MoSold)%>%summarise(.,total = n(), median_p=median(SalePrice))
house_mosold$MoSold <- factor(house_mosold$MoSold, levels=reorder(house_mosold$MoSold, -house_mosold$total))
g4 = ggplot(data=house_mosold, aes(x = MoSold, y = total)) 
plot4 = g4 + geom_bar(aes(fill=median_p), stat = "identity") + 
  geom_text(aes(label = median_p ), size = 3.5,  hjust = 0.5, vjust = -0.3)+
  scale_fill_continuous(guide="colorbar",low ="#56B1F7", high = "#132B43", name= "Median Price($)")+
  xlab("") +
  ylab("Number of Houses Sold")+
  scale_x_discrete(labels=c("Jan", "Feb", "Mar","Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct","Nov", "Dec"))+
  ggtitle("Houses Sale vs. Sold Month")
plot4
```

## House Prices with Different Zones
```{r, echo= F}
plot5_violin = ggplot(data = house_data, aes(x= reorder(MSZoning,SalePrice, median), y =SalePrice)) +
  geom_violin(aes(fill = MSZoning)) + scale_y_continuous(labels = comma) +
  xlab("House Location Zone ")+ 
  ylab("Sale Price ($)")+
  ggtitle("General House Location Zone vs. Sale Price")+
  scale_fill_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854'),
                    name = " ", labels=c("Others", "Floating Village", "Residential High Density",
                                         "Residential Low Density", "Residential Medium Density"))+
  scale_x_discrete(labels=c("Others", "RM", "RH","RL", "FV")) +
  theme(legend.position ="bottom")

plot5_violin
```

## Foundations Matters House Overall Quality
```{r, echo=F}
house_data$OverallQual = as.factor(house_data$OverallQual)
house_overqual= house_data%>%select(.,OverallQual, SalePrice, Foundation)%>%
  group_by(.,OverallQual)
g6 = ggplot(data=house_overqual, aes(x = OverallQual)) 
plot6=g6+geom_bar(aes(fill= Foundation), position = "fill") + 
  theme(legend.position ="bottom") + 
  scale_fill_manual(values = c('#66c2a5','#fc8d62','#8da0cb','#e78ac3','#a6d854','#ffd92f'),
                    name = "Foundation", labels=c("Poured Contrete", "Cinder Block", "Brik&Tile",
                                                  "Slab", "Stone","Wood"))+
  xlab("Overall Quality") +
  ylab("Percent")+
  ggtitle("House Foundation VS. Overall House Quality ")

plot6

```

## Conclusion 

* Popular house style changes with time 
      + Some house style no longer popular and disappear in the market
* It is possible to get a deal in off peak house market season.
* Estimate house price from building structure and house location
* House built year and house overall quality relationship in future analysis

**Thank you!**



