---
title: "Project_1"
output: html_document
---

#Data Visualization On Speed Dating Dataset              
## Dataset Introduction 

***

Data was gathered from participants in experimental speed dating events from 2002-2004. The subjects were drawn from students in graduate and professional schools at Columbia University. Parcipants were recruited through a combination of mass e-mail and fliers posted throughout the campus and handed out by research assistants. During the events, the attendees would have a 4 minute "first date" with every other participant of the opposite sex. At the end  of their 4 minutes, participants were asked if they would like to see their date again. They were also asked to rate their date on six attributes:

a. Attractiveness
b. Sincerity
c. Intelligence
d. Fun
e. Ambition
f. Shared Interests

The dataset also includes questionnaire data gathered from participants at different points in the process. These fields include: demographics, dating habits, self-perception across key attributes, belifs on what others find valuable in a mate, and ifestyle information.

***

## Data Description 

```{r, echo = TRUE, warning = FALSE, message = FALSE}
# Data reading in
setwd("~/Desktop/NYCDSA/Project/Project_1")
library(ggplot2)
library(scales)
library(reshape2)
library(dplyr)
library(ggthemes)
library(grid)
date <- read.csv("Speed Dating Data.csv", 
                 sep = ',', stringsAsFactors = FALSE,
                 header = TRUE)
date1 <- date
dim(date1) 
colnames(date1)
```

```{r, echo = FALSE}
# - - - - - - - - # - - - - - - - - #
#         Multiplot Function        #
# - - - - - - - - # - - - - - - - - #
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)
  
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
  
  numPlots = length(plots)
  
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }
  
  if (numPlots==1) {
    print(plots[[1]])
    
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    
    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

The are 551 people participating in speed dating events, and unique IDs were assigned to these participants. The dataset has 8378 observations and 195 variables, recording all the information for every single dates.

The 195 variables could be divided into 5 categories, based on their backgrounds:
a. Personal information
b. Sports (17 var.)
c. Partner's information
d. Scorecard (9 var.)
e. Others

***

## Data Preprocessing
#### a. Select sunjects in waves without variation

A total of 21 sessions were ran, and 8 have been omitted because there were variations in these dating waves, that might influence people's final decision. 

Among these 13 waves, although 4 of them used different preference scale than the rest, they were included into our sample dataset in order to maintain the general information of the original dataset. 

Waves 5, 12, 13, 14, 18, 19, 20 and 21 were omitted: 

```{r, echo=TRUE}
date1 <- date1[!date1$wave %in% c(5, 12, 13, 14, 18, 19, 20, 21),]  
date1$match <- as.factor(date1$match)
date1$gender <- as.factor(date1$gender)
nrow(date1)                    # 5322 obs
length(unique(date1$iid))      # 348 individuals   
```

***

#### b. Re-scale scores of attributes

In wave 6-9, the preference scale for the 6 attributes ranges from 1-10, while in others the range is 0-100 for each single attribute, but with a sum socre of these 6 attributes equaling to 100. For the variables categorized as scorecard, the 6 attributes were recorded in scale of 1-10. Re-scaling is implemented so that comparison among different attributes would be more observable. Variables scaled from 1-10 were rescaled to 1-100 based on their weights in corresponding sum score of 6 attributes.  

```{r, echo = FALSE}

# - - - - - - - - - - -# - - - - - - - - - - -# - - - - - - - - - - -# - - - - - - - - - - -#
# Re-scale START 
# Total score for these attributesa becomes 100
date1$sum.attr <- date1$attr + date1$sinc  + date1$intel  + date1$fun	+ date1$amb	+ date1$shar
date1$attr <- date1$attr * 100 / date1$sum.attr 
date1$sinc <- date1$sinc * 100 / date1$sum.attr 
date1$intel <- date1$intel * 100 / date1$sum.attr 
date1$fun <- date1$fun * 100 / date1$sum.attr 
date1$amb <- date1$amb * 100 / date1$sum.attr 
date1$shar <- date1$shar * 100 / date1$sum.attr 
#
# Re-scale questions for waves 6, 7, 8 and 9, where the sum scores are not equal to 100
# 4_1
date1$sum.41 <- date1$attr4_1 + date1$sinc4_1  + date1$intel4_1	+
  date1$fun4_1	+ date1$amb4_1	+ date1$shar4_1
date1$attr4_1 <- date1$attr4_1 * 100 / date1$sum.41 
date1$sinc4_1 <- date1$sinc4_1 * 100 / date1$sum.41
date1$intel4_1 <- date1$intel4_1 * 100 / date1$sum.41
date1$fun4_1 <- date1$fun4_1 * 100 / date1$sum.41
date1$amb4_1 <- date1$amb4_1 * 100 / date1$sum.41
date1$shar4_1 <- date1$shar4_1 * 100 / date1$sum.41 
# 3_1
date1$sum.31 <- date1$attr3_1 + date1$sinc3_1  + date1$intel3_1  +
  date1$fun3_1	+ date1$amb3_1
date1$attr3_1 <- date1$attr3_1 * 100 / date1$sum.31 
date1$sinc3_1 <- date1$sinc3_1 * 100 / date1$sum.31
date1$intel3_1 <- date1$intel3_1 * 100 / date1$sum.31
date1$fun3_1 <- date1$fun3_1 * 100 / date1$sum.31
date1$amb3_1 <- date1$amb3_1 * 100 / date1$sum.31
# 5_1
date1$sum.51 <- date1$attr5_1 + date1$sinc5_1  + date1$intel5_1  +
  date1$fun5_1  + date1$amb5_1 
date1$attr5_1 <- date1$attr5_1 * 100 / date1$sum.51 
date1$sinc5_1 <- date1$sinc5_1 * 100 / date1$sum.51
date1$intel5_1 <- date1$intel5_1 * 100 / date1$sum.51
date1$fun5_1 <- date1$fun5_1 * 100 / date1$sum.51
date1$amb5_1 <- date1$amb5_1 * 100 / date1$sum.51
# 4_2
date1$sum.42 <- date1$attr4_2 + date1$sinc4_2  + date1$intel4_2  +
  date1$fun4_2  + date1$amb4_2 + date1$shar4_2
date1$attr4_2 <- date1$attr4_2 * 100 / date1$sum.42 
date1$sinc4_2 <- date1$sinc4_2 * 100 / date1$sum.42
date1$intel4_2 <- date1$intel4_2 * 100 / date1$sum.42
date1$fun4_2 <- date1$fun4_2 * 100 / date1$sum.42
date1$amb4_2 <- date1$amb4_2 * 100 / date1$sum.42
date1$shar4_2 <- date1$shar4_2 * 100 / date1$sum.42
# 3_2
date1$sum.32 <- date1$attr3_2 + date1$sinc3_2  + date1$intel3_2  +
  date1$fun3_2 + date1$amb3_2
date1$attr3_2 <- date1$attr3_2 * 100 / date1$sum.32
date1$sinc3_2 <- date1$sinc3_2 * 100 / date1$sum.32
date1$intel3_2 <- date1$intel3_2 * 100 / date1$sum.32
date1$fun3_2 <- date1$fun3_2 * 100 / date1$sum.32
date1$amb3_2 <- date1$amb3_2 * 100 / date1$sum.32
# 5_2
date1$sum.52 <- date1$attr5_2 + date1$sinc5_2  + date1$intel5_2  +
  date1$fun5_2 +date1$amb5_2
date1$attr5_2 <- date1$attr5_2 * 100 / date1$sum.52
date1$sinc5_2 <- date1$sinc5_2 * 100 / date1$sum.52
date1$intel5_2 <- date1$intel5_2 * 100 / date1$sum.52
date1$fun5_2 <- date1$fun5_2 * 100 / date1$sum.52
date1$amb5_2 <- date1$amb5_2 * 100 / date1$sum.52
# 4_3
date1$sum.43 <- date1$attr4_3 + date1$sinc4_3  + date1$intel4_3  +
  date1$fun4_3  + date1$amb4_3 + date1$shar4_3
date1$attr4_3 <- date1$attr4_3 * 100 / date1$sum.43 
date1$sinc4_3 <- date1$sinc4_3 * 100 / date1$sum.43
date1$intel4_3 <- date1$intel4_3 * 100 / date1$sum.43
date1$fun4_3 <- date1$fun4_3 * 100 / date1$sum.43
date1$amb4_3 <- date1$amb4_3 * 100 / date1$sum.43
date1$shar4_3 <- date1$shar4_3 * 100 / date1$sum.43
# 3_3
date1$sum.33 <- date1$attr3_3 + date1$sinc3_3  + date1$intel3_3  +
  date1$fun3_3 + date1$amb3_3
date1$attr3_3 <- date1$attr3_3 * 100 / date1$sum.33
date1$sinc3_3 <- date1$sinc3_3 * 100 / date1$sum.33
date1$intel3_3 <- date1$intel3_3 * 100 / date1$sum.33
date1$fun3_3 <- date1$fun3_3 * 100 / date1$sum.33
date1$amb3_3 <- date1$amb3_3 * 100 / date1$sum.33
# 2_3
date1$sum.23 <- date1$attr2_3 + date1$sinc2_3  + date1$intel2_3  +
  date1$fun2_3  + date1$amb2_3 + date1$shar2_3
date1$attr2_3 <- date1$attr2_3 * 100 / date1$sum.23 
date1$sinc2_3 <- date1$sinc2_3 * 100 / date1$sum.23
date1$intel2_3 <- date1$intel2_3 * 100 / date1$sum.23
date1$fun2_3 <- date1$fun2_3 * 100 / date1$sum.23
date1$amb2_3 <- date1$amb2_3 * 100 / date1$sum.23
date1$shar2_3 <- date1$shar2_3 * 100 / date1$sum.23
# 5_3
date1$sum.53 <- date1$attr5_3 + date1$sinc5_3  + date1$intel5_3  +
  date1$fun5_3  + date1$amb5_3 
date1$attr5_3 <- date1$attr5_3 * 100 / date1$sum.53 
date1$sinc5_3 <- date1$sinc5_3 * 100 / date1$sum.53
date1$intel5_3 <- date1$intel5_3 * 100 / date1$sum.53
date1$fun5_3 <- date1$fun5_3 * 100 / date1$sum.53
date1$amb5_3 <- date1$amb5_3 * 100 / date1$sum.53

# Re-scale DONE
# - - - - - - - - - - -# - - - - - - - - - - -# - - - - - - - - - - -# - - - - - - - - - - -#

```

***

##Pre-Research Questions: 
#####a. What are the least desirable attributes in a male partner? Does this differ for female partners?
#####b. Will being physically attractive help you get more dates?



```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.width=20, fig.height=10}
# First Level EDA
get.date <- date1 
get.date$match <- as.numeric(get.date$match) 
get.date$match <- ifelse(get.date$match == 1, 0, 1) # done dealing with the factor variable
get.date1 <- get.date %>%
  group_by(iid) %>%
  summarize(MatchScore = sum(match))
get.date1 <- as.data.frame(table(get.date1$MatchScore))

# The plot of the distribution of matching number
eda.1 <- ggplot(get.date1, aes(x = Var1, y = Freq)) +
  geom_bar(aes(y = Freq), stat = 'identity') +
  labs(x = 'Number of total dating match', y = 'Number of participants') +
  theme_bw() +
  ggtitle('A. Distribution of matching number') +
  theme(plot.title=element_text(face="bold", size = rel(1.5)))

# The plot of the distribution of age among participants
eda.2.data <- date1 %>% 
  select(iid, gender, age)
eda.2.data <- distinct(eda.2.data)
eda.2.data$gender <- as.factor(eda.2.data$gender)
levels(eda.2.data$gender) <- c('Female','Male')
#
eda.2.plt <- ggplot(eda.2.data) +
  geom_density(aes(x = age, y = ..count.., color = gender), lwd = 1) 
eda.2.plt1 <- ggplot_build(eda.2.plt)
eda.2.dat2 <- data.frame(x = eda.2.plt1$data[[1]]$x[eda.2.plt1$data[[1]]$group==1],
                  ymin=eda.2.plt1$data[[1]]$y[eda.2.plt1$data[[1]]$group==1],
                  ymax=eda.2.plt1$data[[1]]$y[eda.2.plt1$data[[1]]$group==2])
eda.2.dat2$ymax[eda.2.dat2$ymax > eda.2.dat2$ymin] <- eda.2.dat2$ymin[eda.2.dat2$ymax > eda.2.dat2$ymin]

eda.2.dat2.2 <- data.frame(x = eda.2.plt1$data[[1]]$x[eda.2.plt1$data[[1]]$group==1],
                  ymin=eda.2.plt1$data[[1]]$y[eda.2.plt1$data[[1]]$group==1],
                  ymax=eda.2.plt1$data[[1]]$y[eda.2.plt1$data[[1]]$group==2])
eda.2.dat2.2$ymax[eda.2.dat2.2$ymax < eda.2.dat2.2$ymin] <- eda.2.dat2.2$ymin[eda.2.dat2.2$ymax < eda.2.dat2.2$ymin]
  
eda.2 <- eda.2.plt +
  geom_ribbon(data = eda.2.dat2, aes(x = x, ymin = ymin, ymax = ymax), 
              fill = 'red3', alpha = 0.3) + 
  geom_ribbon(data = eda.2.dat2.2, aes(x = x, ymin = ymin, ymax = ymax), 
              fill = '#33CCCC', alpha = 0.3) +
  labs(x = 'Age', y = 'Density') +
  theme_bw() +
  ggtitle('B. Distribution of age among participants') +
  theme(plot.title=element_text(face="bold", size = rel(1.5))) 

# Plot of the participants' fields of study
eda.3.data <- distinct(date1[complete.cases(date1$field_cd), c('iid','field_cd')])
eda.3.data$field_cd <- as.factor(eda.3.data$field_cd)
eda.3 <- 
  ggplot(eda.3.data, aes(x = field_cd)) +
  geom_bar() +
  labs(x = 'Field of Study', y = 'Number of Participants') +
  scale_x_discrete(breaks = c(1:16),
                      labels = c('LAW', 'MAT','PSY','MED','ENGR','WRI','HIST',
                                 'FIN','EDU','SCI','Soc.W','UND','POL',
                                 'FIL','ART','LAN')) +
  theme(axis.text.x = element_text(face="italic", color="red4", 
                                   size=10, angle=45, vjust = -0.1)) +
  ggtitle("B. Distribution of participants' field of study") +
  theme_bw() +
  theme(plot.title=element_text(face="bold", size = rel(1.5)))

# Plot of the participants' expectation of happiness aobut their speed-dating partners.
eda.4.data <- distinct(date1[complete.cases(date1$exphappy), c('iid','gender', 'exphappy')])
levels(eda.4.data$gender) <- c('Female','Male')
eda.4.data$gender <- as.factor(eda.4.data$gender)
eda.4.plt <- ggplot(eda.4.data) +
  geom_density(aes(x =  exphappy, y = ..count.., color = gender), lwd = 1) 
eda.4.plt1 <- ggplot_build(eda.4.plt)
eda.4.dat4 <- data.frame(x = eda.4.plt1$data[[1]]$x[eda.4.plt1$data[[1]]$group==1],
                  ymin=eda.4.plt1$data[[1]]$y[eda.4.plt1$data[[1]]$group==1],
                  ymax=eda.4.plt1$data[[1]]$y[eda.4.plt1$data[[1]]$group==2])
eda.4.dat4$ymax[eda.4.dat4$ymax > eda.4.dat4$ymin] <- eda.4.dat4$ymin[eda.4.dat4$ymax > eda.4.dat4$ymin]

eda.4.dat4.4 <- data.frame(x = eda.4.plt1$data[[1]]$x[eda.4.plt1$data[[1]]$group==1],
                  ymin=eda.4.plt1$data[[1]]$y[eda.4.plt1$data[[1]]$group==1],
                  ymax=eda.4.plt1$data[[1]]$y[eda.4.plt1$data[[1]]$group==2])
eda.4.dat4.4$ymax[eda.4.dat4.4$ymax < eda.4.dat4.4$ymin] <- eda.4.dat4.4$ymin[eda.4.dat4.4$ymax < eda.4.dat4.4$ymin]
  
eda.4 <- eda.4.plt +
  geom_ribbon(data = eda.4.dat4, aes(x = x, ymin = ymin, ymax = ymax), 
              fill = 'red3', alpha = 0.3) + 
  geom_ribbon(data = eda.4.dat4.4, aes(x = x, ymin = ymin, ymax = ymax), 
              fill = '#33CCCC', alpha = 0.3) +
  labs(x = 'Happiness Expecation (Scale of 1-10)', y = 'Number of Participants') +
  ggtitle("D. Distribution of participants' dating happiness expectation") +
  theme_bw() +
  theme(plot.title=element_text(face="bold", size = rel(1.5)))

```

```{r, echo = FALSE}
attr.sur6 <- c('attr','sinc','intel','fun','amb','shar') # 6 attributes
attr.sur5 <- c('attr','sinc','intel','fun','amb') # 5 attributes
date1.1 <- date1
#
q2.data <- date1.1 %>%
  select(gender, contains('1_'), contains('2_'), 
         contains('3_'), contains('4_'), contains('5_'), 
         contains('7_')) 
q2.data <- q2.data %>%
  group_by(gender) %>%
  summarise_each(funs(mean(., na.rm = TRUE)))
s1 <- q2.data %>%
  select(gender, contains('_1'))
s2 <- q2.data %>%
  select(gender, contains('_2'))
s3 <- q2.data %>%
  select(gender, contains('_3'))
s1$survey <- 1
s2$survey <- 2
s3$survey <- 3
colnames(s1) <- c('gender', 
                  paste0(attr.sur6, '1'),
                  paste0(attr.sur6, '2'),
                  paste0(attr.sur5, '3'),
                  paste0(attr.sur6, '4'),
                  paste0(attr.sur5, '5'), 
                  'survey')
colnames(s2) <- c(colnames(s1)[-30], 
                  paste0(attr.sur6, '7'), 'survey')
colnames(s3) <- colnames(s2)
s1.1 <- melt(s1, id.var = c('gender', 'survey'))
s2.2 <- melt(s2, id.var = c('gender', 'survey'))
s3.3 <- melt(s3, id.var = c('gender', 'survey'))
s123 <- rbind(s1.1, s2.2, s3.3) # containing all survey questions answers labeled by survey number and gender
#
q1 <- s123[s123$variable %in% paste0(attr.sur6, '1'),] # 1
q1$survey <- as.factor(q1$survey)
levels(q1$variable)[1:6] <- c('Attractive','Sincere','Intelligent','Fun','Ambitious','Shared Interest')
levels(q1$gender) <- c('Female','Male')
q2 <- s123[s123$variable %in% paste0(attr.sur6, '2'),] # 2
q2$survey <- as.factor(q2$survey)
levels(q2$variable)[7:12] <- c('Attractive','Sincere','Intelligent','Fun','Ambitious','Shared Interest')
levels(q2$gender) <- c('Female','Male')
q3 <- s123[s123$variable %in% paste0(attr.sur5, '3'),] # 3
q3$survey <- as.factor(q3$survey)
levels(q3$variable)[13:17] <- c('Attractive','Sincere','Intelligent','Fun','Ambitious')
levels(q3$gender) <- c('Female','Male')
q4 <- s123[s123$variable %in% paste0(attr.sur6, '4'),] # 4
q4$survey <- as.factor(q4$survey)
levels(q4$variable)[18:23] <- c('Attractive','Sincere','Intelligent','Fun','Ambitious','Shared Interest')
levels(q4$gender) <- c('Female','Male')
q5 <- s123[s123$variable %in% paste0(attr.sur5, '5'),] # 5
q5$survey <- as.factor(q5$survey)
levels(q5$variable)[24:28] <- c('Attractive','Sincere','Intelligent','Fun','Ambitious')
levels(q5$gender) <- c('Female','Male')
q7 <- s123[s123$variable %in% paste0(attr.sur6, '7'),] # 7
q7$survey <- as.factor(q7$survey)
levels(q7$variable)[29:34] <- c('Attractive','Sincere','Intelligent','Fun','Ambitious','Shared Interest')
levels(q7$gender) <- c('Female','Male')
```

```{r, echo = FALSE, warning = FALSE, message = FALSE}
# Group-by sex
sex <- date1.1 %>% 
  group_by(gender) %>%
  summarize(m.attr1_1 = mean(attr1_1, na.rm = TRUE),
            m.sinc1_1 = mean(sinc1_1, na.rm = TRUE),
            m.intel1_1 = mean(intel1_1, na.rm = TRUE),
            m.fun1_1 = mean(fun1_1, na.rm = TRUE),
            m.amb1_1 = mean(amb1_1, na.rm = TRUE),
            m.shar1_1 = mean(shar1_1, na.rm = TRUE))
sex <- melt(sex, id.var = 'gender')


sc.var <- c('attr','sinc','intel','fun','amb')
sc <- date1[, c('gender', sc.var)]
sc <- sc %>% 
  group_by(gender) %>%
  summarise_each(funs(mean(., na.rm = TRUE)))
sc$survey <- 'Scorecard Rate'
colnames(sc) <- c('gender','Attractive','Sincere','Intelligent',
                  'Fun','Ambitious','survey')
levels(sc$gender) <- c('Female','Male')
sc.melt <- melt(sc, id.var = c('gender','survey'))
sc.q3 <- rbind(sc.melt, q3[q3$survey == 1,])
colnames(sc.q3)[2] <- 'Source_of_Scores'
sc.q3[sc.q3[,2]=="1", 2] <- 'Question 3 in Survey 1'
```

```{r, echo = FALSE, fig.width=20, fig.height=10, warning = FALSE, fig.width=20, fig.height=10}
p1 <- ggplot(sex, aes(x = variable, y = value, fill = gender))+
  geom_bar(aes(group = gender), stat = 'identity', position = 'dodge') +
  geom_text(aes(label=round(value), y=value), 
            stat="identity",vjust=-0.5, size=4, 
            position = position_dodge(width=0.9)) +
  labs(x = 'Attributes', y = 'Score') +
  ggtitle('C. Attribute Preference For Participants In Different Genders') +
  scale_fill_discrete(breaks = c(0, 1),
                        labels = c('Female', 'Male'),
                        guide = guide_legend(direction = "vertical", 
                                             title.position = "top")) +
  scale_x_discrete(labels = c('Attractive', 'Sincere', "Intelligent",'Fun',
                              'Ambitious', 'Shared Interests'))  +
  theme(plot.title=element_text(face="bold", size = rel(1.5)))

p3 <- ggplot(sc.q3, aes(x = variable, y = value, fill = Source_of_Scores, group = gender)) +
  geom_bar(aes(group = Source_of_Scores), stat = 'identity', position = 'dodge') + 
  facet_grid(gender~.) + 
  scale_y_continuous(limits = c(0, 25)) +
  labs(x = 'Attributes', y = 'Score') +
  ggtitle('F. Mean Scores From Participants vs. Mean Scores From Their Partners') +
  scale_x_discrete(labels = c('Attractive', 'Sincere', "Intelligent",'Fun',
                              'Ambitious', 'Shared Interests')) +
  theme(plot.title=element_text(face="bold", size = rel(1.5)))
```

```{r, echo = FALSE, message = FALSE, warning=FALSE, fig.width=20, fig.height=10}
multiplot(eda.1, eda.2, cols = 2)
multiplot(p1, eda.4, cols = 2)
#multiplot(p1, p3, cols = 2)
```

*Figure 1. Brief exploratory data analysis for chosen attributes based on the research questions*  

***

* Observations from these plots:
     a. **Distribution of Number of Matches** *(Upper-Left)*
     + Around 60 participants received no dating match.
     b. **Density of Age** *(Upper-Right)*
     + The peak of the density plot for male participants located at the age greater than that for female participants. 
     c. **Attributes' Preferences** *(Bottom-Left)*
     + Female participants put greater weight on non-physical attributes, while male participants place more value on physical attractiveness.
     d. **Density of Happiness Expectation of Speed-Dating** *(Bottom-Right)*
     + Males might expect higher happiness for a date  than females do. 

***

##How were participants' answer on same survey questions varying at different stage of dating?


```{r, echo = FALSE, fig.width=20, fig.height=8, warning = FALSE}
# - - - - - - - - - - - - - - 
#
# ggplot setups
bar <- geom_bar(aes(group = variable), stat = 'identity',position = 'dodge')
themes <- theme(plot.title=element_text(face="bold", size = rel(1.2)),
        axis.title.x=element_text(size=10, face='bold'),
        axis.title.y=element_text(size=10, face='bold'))
labs <- labs(x = 'Survey', y = 'Score')
face <- facet_grid(gender~variable)
lty <- geom_hline(aes(yintercept = 20), linetype = 3)
ylimit <- scale_y_continuous(limits = c(0, 40))
#

q1.plt <- ggplot(q1, aes(x = survey, y = value, fill = variable, group = gender)) +
  ggtitle('1.What you look for in the opposite sex?') + guides(fill = FALSE) + 
  bar + themes + labs + face + lty + ylimit +
  annotate("text", x = Inf, y = -Inf, label = "Q.1",
           hjust=1.1, vjust=-1.1, col="black", cex=6,
           fontface = "bold", alpha = 0.8, size = 10)

q2.plt <- ggplot(q2, aes(x = survey, y = value, fill = variable, group = gender)) +
  ggtitle('2.What do you think the opposite sex looks for in a date') + guides(fill = FALSE) +
  bar + themes + labs + face + lty + ylimit +
  annotate("text", x = Inf, y = -Inf, label = "Q.2",
           hjust=1.1, vjust=-1.1, col="black", cex=6,
           fontface = "bold", alpha = 0.8, size = 10)

q3.plt <- ggplot(q3, aes(x = survey, y = value, fill = variable, group = gender)) +
  ggtitle('3.How do you think you measure up?') + guides(fill = FALSE) +
  bar + themes + labs + face + lty + ylimit +
  annotate("text", x = Inf, y = -Inf, label = "Q.3",
           hjust=1.1, vjust=-1.1, col="black", cex=6,
           fontface = "bold", alpha = 0.8, size = 10)

q4.plt <- ggplot(q4, aes(x = survey, y = value, fill = variable, group = gender)) +
  ggtitle('4.What you think MOST of your fellow men/women \n look for in the opposite sex?') +
  guides(fill = FALSE) + bar + themes + labs + face + lty + ylimit +
  annotate("text", x = Inf, y = -Inf, label = "Q.4",
           hjust=1.1, vjust=-1.1, col="black", cex=6,
           fontface = "bold", alpha = 0.8, size = 10)

q5.plt <- ggplot(q5, aes(x = survey, y = value, fill = variable, group = gender)) +
  ggtitle('5.How do you think others perceive you?') + guides(fill = FALSE) +
  bar + themes + labs + face + lty + ylimit +
  annotate("text", x = Inf, y = -Inf, label = "Q.5",
           hjust=1.1, vjust=-1.1, col="black", cex=6,
           fontface = "bold", alpha = 0.8, size = 10)

q7.plt <- ggplot(q7, aes(x = survey, y = value, fill = variable, group = gender)) +
  ggtitle('6.Distribute the points that best reflects the actual \n importance of these attributes in your decisions.') +
  guides(fill = FALSE) + bar + themes + labs + face + lty + ylimit +
  annotate("text", x = Inf, y = -Inf, label = "Q.6",
           hjust=1.1, vjust=-1.1, col="black", cex=6,
           fontface = "bold", alpha = 0.8, size = 10)

multiplot(q1.plt, q2.plt, cols = 2)
multiplot(q3.plt, q7.plt, cols = 2)
```
    
*Figure 2. The answers for question in the surveys. Surveys were taken at different stage of dating in chronological order.*

***

* Inferences:
     a. **What you look for in the opposite sex?**
     + Males weight more on physical attractiveness.
     + Females' weight on different attributes are more balanced than males.
     b. **What do you think the opposite sex looks for in a date?**
     + Males think females would weight more on physical attractiveness than on other attributes.
     + Females think males would weight more on physical attractiveness than on other attribute. 
     c. **How do you think you measure up?**
     + Males evaluate their attributes almost evenly.
     + Femals evaluate their attributes almost evenly.
     d. **Distribute the points that best reflects the actual importance of these attributes in your decisions.**
     + Males weight more on physical attractiveness than females do.
     + Females considers on physical attractiveness more than on any other attributes. 


***

##Will being physically attractive help you get more dates?

```{r, echo = FALSE, fig.width=20, fig.height=10, warning = FALSE, fig.width=20, fig.height=10}
q66 <- date1 %>%
  select(iid, pid, gender, match, match_es, date, attr3_1)
q66$match <- ifelse(q66$match == 1, 1, 0)
q66$date <- as.factor(q66$date)
#q66 <- q66[-which(q66$match_es==18),] # The subject with iid 220 is omitted because he might not take the survey seriously. 
#date1[date1$iid == 220, 'match']
q66.box.attr <- date1 %>% select(iid, gender, match_es, date, attr3_1)
q66.box.attr <- distinct(q66.box.attr)
num.match <- q66 %>% group_by(iid) %>% summarize(n.Match = sum(match))
tile.p <- merge(num.match, q66.box.attr, by = 'iid')
tile.p <- tile.p %>%
  group_by(n.Match, date) %>%
  summarize(n = n())
tile.p$percentage <- tile.p$n/nrow(tile.p) * 100

tile.date <- ggplot(tile.p, aes(date, n.Match, fill = percentage)) +
  geom_tile(show.legend = FALSE) + 
  theme_bw() +
  labs(x = 'Number of Dating Match', 
       y = 'Number of Dates (per week)') +
  ggtitle('Out-Date Frequency vs. Number of Match After Sp.Dt.') +
  theme(plot.title=element_text(face="bold", size = rel(1.2)),
        axis.title.x=element_text(size=10, face='bold'),
        axis.title.y=element_text(size=10, face='bold'))
#____________________________#____________________________#____________________________#
attr_o <- date1 %>%
  select(iid, pid, attr_o, like, prob, met) 
# Create unique identifier
attr_o$iid <- as.character(attr_o$iid)    # For attr_o
attr_o$pid <- as.character(attr_o$pid)
attr_o$unique <- paste0(attr_o$pid, '+', attr_o$iid)
q66$iid <- as.character(q66$iid)    # For q66
q66$pid <- as.character(q66$pid)
q66$unique <- paste0(q66$pid, '+', q66$iid)
# merge
att <- merge(q66, attr_o, by = 'unique', all = FALSE)
att$iid.y <- NULL
att$pid.x <- NULL
att$pid.y <- NULL
att$iid <- att$iid.x
att$iid.x <- NULL
att$met <- ifelse(att$met == 1, 1, 0)
att$met <- as.factor(att$met)
att$iid <- as.numeric(att$iid)
att$gender <- ifelse(att$gender == 1, 'Male','Female')
att <- att %>% arrange(iid)

att$attr_o <- round(att$attr_o)
# summary(att$attr_o)
att$attr_oo <- cut(att$attr_o, c(0, 5, 6, 10), labels = c('Low','Medium', 'High'))
att.box <- att[complete.cases(att$attr_oo), ]
phy.attr <- ggplot(att.box, aes(x = attr_oo, y = like)) +
  geom_boxplot(aes(fill = gender)) + 
  theme_igray() + 
  facet_grid(~gender) +
  labs(x = "Participants' Physical Attractiveness (Scored by dating partners)", 
       y = "How much do you like this person? (Scored by dating partners)") +
  ggtitle("Wil Physical Attractiveness Affect Partners' liking degree?") +
  theme(plot.title=element_text(face="bold", size = rel(1.2)),
        axis.title.x=element_text(size=10, face='bold'),
        axis.title.y=element_text(size=10, face='bold'))
# Another dataset

mean.att <- att.box %>%
  select(iid, gender, attr_o) %>%
  group_by(iid, gender) %>%
  summarize(mean.attr = mean(attr_o, na.rm = TRUE))
att.match <- merge(mean.att, num.match, by = 'iid')
att.match$n.Match <- as.factor(att.match$n.Match)
att.match.plt <- ggplot(att.match, aes(x = n.Match, y = mean.attr)) +
  geom_boxplot(aes(fill = gender)) +
  facet_grid(~gender) +
  theme_bw() + theme_igray() +
  labs(x = "Number Of Total Dating Match Obtained After Speed-Dating", 
       y = "Physical Attractiveness Scored By Partners") +
  ggtitle("Physical Attractiveness' Affects On Number Of Date Obtained") +
  theme(plot.title=element_text(face="bold", size = rel(1.2)),
        axis.title.x=element_text(size=10, face='bold'),
        axis.title.y=element_text(size=10, face='bold'))

match.att <- att %>% select(iid, gender, met, attr_o)
match.att <- match.att[complete.cases(match.att$met),]
match.att$met <- ifelse(match.att$met == 0, 'No', 'Yes')
met.yn <- ggplot(match.att, aes(x = met, y = attr_o))+
  geom_boxplot() +
  theme_bw() + theme_igray() +
  labs(x = "Have the participants and their dating partners met before?", 
       y = "Physical Attractiveness Scored By Partners") +
  ggtitle("Physical Attractiveness For Participants Met (Yes/No) Before") +
  theme(plot.title=element_text(face="bold", size = rel(1.2)),
        axis.title.x=element_text(size=10, face='bold'),
        axis.title.y=element_text(size=10, face='bold'))



phy.attr
att.match.plt
met.yn
```

***

* Observations:
     + Males, with a high physical attractiveness, gained higher mean liking degree than males in other groups. 
     + Females gained similar liking degree from their partners, despite the difference in physical attractiveness.
     + The sensitivity towards physical attractiveness of females is stronger than that of males. 
     + The more the number of total match obtained, the higher the mean of participants' physical attractiveness. 
     
***

##Inferences:
a. Physical attractiveness is probably the most considered and valued attribute among those six attributes: attractiveness, sincerity, intelligence, fun, ambition, and shared interests.

b. Even though males valued physical attractiveness, females seem to have high sensitivity on this attribute than do males.

***

##Further Questions:
a. Are there any common attributes that can be found from participants received no dating match?
b. Is ambition another influencing attribute in speed-dating?
c. When choosing dating partner, how is age and race being weighted?

***

##Limitations:
a. The sample size is too small and the survey subjects were collected from Columbia University; therefore bias might be introduced.
b. The survey questions were not comprehensive.
c. Some key features of participants were not recorded. (i.e. number of ex-, length of the longest relationship they had, etc.)

     





