---
title: "Lending Club Loan Visualization"
output:
  ioslides_presentation:
    smaller: yes
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
library(ggplot2)
library(scales) 
library(dplyr)
library(tidyr)
library(tm)
library(SnowballC)
library(wordcloud)

wd = '~/R files/lendingClub'

setwd(wd)

l2013 <- read.table("proc_Loans2013.csv", sep=";", quote="\"", header=T, comment.char="", encoding="UTF-8", fill=F)

l <- read.table("proc_Loans40k.csv", sep=";", quote="\"", header=T, comment.char="", encoding="UTF-8", fill=F)

l$Trm.issue_d = as.Date(l$Trm.issue_d)
```

## Introduction
### Lending Club is a leading peer-to-peer lender  

###How it works:  
- Lending Club uses potential borrowers' credit history to estimate credit risk (the risk of not being paid back)  
- This credit risk estimate is used to set the interest rate -- higher risk, higher return  
- Lenders (e.g., you and I) then choose which loans to fund at the LC-prescribed rate  

#### Analyzed historical lending data available on the LC website ![](LC_img.tiff)

## Why do people use Lending Club?

```{r}
c <- sample_frac(l2013, 0.2)

c <- Corpus(VectorSource(tolower(c$Trm.title)))

c <- tm_map(c, PlainTextDocument)

c <- tm_map(c, removePunctuation)

c <- tm_map(c, removeWords, c("borrow", "loan", "will", "payment", stopwords('english')))

wordcloud(c, max.words = 100, random.order = FALSE)
```
<div class=centered>
Most borrowing is used to repay existing debt
</div>
## Loan Grade Drives Interest Rate
```{r fig.height=4}
ggplot(l, aes(x=Trm.issue_d, col=Trm.grade, y=Trm.int_rate))+
  geom_point() +
  theme_minimal() +
  ggtitle("Interest Rate by Loan Grade") + 
  scale_color_brewer(palette="Spectral", name="Grade", direction=-1)+
  scale_x_date(date_labels = "%Y", date_breaks = "1 year", name="Issue Date") +
  scale_y_continuous(labels=percent_format(), name="Interest Rate") 
```

- Rates have been stable for high grade loans, but trending higher for others  
- Getting bumped down a notch or two can mean 5-10% difference in interest rate (~\$750--1500 on a \$10k loan)  

## Credit Profile Trends
```{r}
# box plots showing reduced dependence on FICO score and increased number of outliers

ggplot(l,aes(x=Trm.grade, y=Prof.fico_avg)) + 
  geom_boxplot() + 
  facet_wrap(~Trm.issue_Y) + 
  ylim(650,850) + 
  theme_minimal() +
  xlab("Loan Grade") +
  ylab("FICO Score") +
  ggtitle("Loan Grade by FICO Score Trend")
```

Less dependence on FICO scores over the years

## Smarter Lending at the Lending Club?
```{r fig.height=4}

ggplot(subset(l, Trm.issue_Y==2016 |Trm.issue_Y==2007),aes(x=Trm.grade, y=Prof.fico_avg)) + 
  geom_jitter(shape=1, mapping=aes(color=outlierFlg)) + 
  facet_wrap(~Trm.issue_Y) + 
  ylim(650,850) + 
  theme_minimal() +
  xlab("Loan Grade") +
  ylab("FICO Score") +
  ggtitle("") + 
  scale_color_manual(name="", labels=c("Not Outlier", "Outlier"), values=(c(muted("blue"),muted("red")))) +
  geom_boxplot(data=subset(l, (Trm.issue_Y==2016 |Trm.issue_Y==2007)), fill=NA, color="black", outlier.color=NA)
```
Breaking the FICO mould:  
- Borrowers with lower FICO scores paying low interest rates  
- Borrowers with excellent FICO scores paying high interest rates ("outliers")  
- Is Lending Club making a mistake?  

## Outliers Belong in their Buckets
```{r}
t <- l2013 %>% group_by(outlierFlg,Trm.grade, Stat.ChargeOffFlg) %>% 
  summarize(Stat.loan_amnt = sum(Stat.loan_amnt)) 
t <- t %>% group_by(outlierFlg,Trm.grade) %>% 
  mutate(pctOfNotional=Stat.loan_amnt / sum(Stat.loan_amnt)) %>% ungroup

ggplot(subset(t, Stat.ChargeOffFlg==T), aes(x=Trm.grade, y=pctOfNotional, fill=outlierFlg)) +
  geom_bar(stat="identity", position="dodge")+
  theme_minimal() +
  ggtitle("Default Rates 2013") + 
  scale_fill_manual(name="Oulier", labels=c("Not Outlier","Outlier"),   values=c(muted("blue"),muted("red")))+
  scale_y_continuous(labels=percent_format(), name="Percent of Notional Defaulted") +
  xlab("Loan Grade")
```
  
Default rates for outliers better than their peers, but they can't be bumped up to a higher grade

## What's Bad about the Outliers?
```{r fig.height=3.5}

# numerics for faceting 
t <- l2013 %>% select(Trm.issue_Y, 
                  hypGrade, 
                  outlierFlg, 
                  Loan_Amount=Stat.loan_amntpctRank, 
                  Debt_Inc_ratio=Prof.dtipctRank, 
                  Erlst_crd_line=Prof.earliest_cr_line_YpctRank,
                  Mth_snc_rcnt_acct = Prof.mo_sin_rcnt_tlpctRank,
                  Mortgage_cnt = Prof.mort_accpctRank,
                  Mth_snc_derog_rpt = Prof.mths_since_last_major_derogpctRank,
                  Mth_snc_last_rec = Prof.mths_since_last_recordpctRank,
                  Mths_snc_rcnt_card = Prof.mths_since_recent_bcpctRank,
                  Mths_snc_card_delinq = Prof.mths_since_recent_bc_dlqpctRank,
                  Mths_snc_rcnt_inq = Prof.mths_since_recent_inqpctRank,
                  Mths_snc_revol_delinq=Prof.mths_since_recent_revol_delinqpctRank,
                  Loan_income_ratio=Prof.LCLoanIncRatiopctRank) %>%
    gather(., key=attrib_name, value=attrib_value, -Trm.issue_Y, -hypGrade, -outlierFlg)

ggplot(t, aes(x=outlierFlg, y=attrib_value, fill=outlierFlg )) + 
  geom_violin() +
  facet_wrap(~attrib_name) +
  theme_minimal() +
  ggtitle("Loan Features 2013") +
  scale_x_discrete(labels=element_blank(), name="") +
  scale_y_continuous(name="Percentile", labels=percent_format()) +
  scale_fill_manual(values=c(muted("blue"),muted("red")), name="", labels=c("Not outlier", "Outlier"))
```
  
Slight bias towards:  
- Higher nominal loan amounts  
- Higher loan amounts as proportion of income  
- Increased indicators of economic distress (recently opened a new credit line, recent credit inquiries)

## Conclusion

#### Takeaways
- FICO still matters --- lend to them instead of lower scored borrowers in the same loan grade
- Specific driver of getting bumped down a grade not determined

#### Next Steps
- Bootcamp 6's Kaggle winners used training data that had additional useful information about the borrowers that is not in the publically available data --- what is LC keeping from us?  
- Multivariate analysis --- possibly a set of themes for getting put in the outlier credit doghouse
- Integration of denied loan request data
- Analysis of low FICO, high LC grade loans