---
title: "EDA"
author: "Regan Yee"
date: "November 16, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(Hmisc)
library(corrplot)
library(DT)
library(caret)
setwd("~/Documents/Project/4-Kaggle/")
train = readRDS(file = "train.RDS")
test = readRDS(file = "test.RDS")

# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

## Allstate Kaggle

**Allstate Kaggle EDA:** <https://www.kaggle.com/c/allstate-claims-severity>
Training set is 188318 rows 
131 variables (without loss variable)
id + 116 categorical + 14 continuous variables


Test set is 125546 rows
Total data set is 313864 items.

Possible Insights:
- All the categorical values are between 0 and 1

## Summaries

```{r summary}
summary(train)
summary(test)
```

## Plot of Losses Distribution

```{r plot_losses, echo=FALSE, warning=FALSE}
g = ggplot(train, mapping = aes(x=loss))
g + geom_histogram(binwidth = 30)

g = ggplot(train, mapping = aes(x=log(loss)))
g + geom_histogram(binwidth = .1)
```

## Plot Correlations of Categorical Features
```{r binmap, echo=FALSE}
## 116 categorical variables:2-116
names(train)
xplots = list()
for(i in 2:116){
  g2 = ggplot(train, mapping = aes(x=id, y=log(loss), fill=train[,i]))
  g2_plot = g2 + 
            geom_bin2d(bins=100) + 
            xlab(names(train)[i])  + 
            ggtitle(paste0("Binmap of ", names(train)[i]))+
            scale_fill_discrete(name = names(train)[i])
  print(g2_plot)
}

g2 = ggplot(train, mapping = aes(x=id, y=log(loss), fill=cat1))

g2_plot = g2 + geom_bin2d(bins=100) + facet_grid(~ train$cat1)
            xlab(names(train)[i])  + 
            ggtitle(paste0("Binmap of ", names(train)[i]))+
            scale_fill_discrete(name = names(train)[i])
```

## cat116
```{r cat116}
cat116_analysis = train %>% 
      select(cat116,loss) %>% 
      group_by(cat116) %>% 
      summarise(count=n()) %>% 
      arrange(desc(count))
datatable(cat116_analysis)

cat116_analysis2 = train %>% 
      select(cat116,loss) %>% 
      group_by(cat116) %>% 
      summarise(mean=mean(loss)) %>% 
      arrange(desc(mean))
datatable(cat116_analysis2)
```

## cat112
```{r cat112}
## 51 Levels
table(train$cat112)
```

## Correlation Plot
```{r corr}
M = cor(train[,118:132], method="pearson")
corrplot.mixed(M, upper = "circle", order="hclust")
```


## Preprocess Predictors with near zero-variance
preProc = preProcess(train, method = "nzv")
preProc

## ANOVA
```{r}
summary(aov(train$cont1 ~ train$cont2))
```

unique(cat116_analysis2) = train %>% 
      select(cat116,loss)