---
title: "Web Scraping Project"
author: "Yisong Tao"
date: "11/13/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Web scraping MyAnimeList.net 

MyAnimelist.net scraped reviews and info of 5953 animes.  
1. web scraping  
2. reviews: EDA  
3. anime info: network graph eda  
4. predict anime score  



1. web scraping
Done with scrapy. Scraped 2 parts of MyAnimelist.net: anime reviews and anime info.
89244 reviews, 5953 anime info downloaded.
Lesson learned:  
- clean up data while scraping, 
- store data in correct format -- Json vs. csv -> list vs. data.frame.


2. reviews EDA
```{r load data and library, warning=F, message=F}
library(ggplot2)
library(dplyr)
AnimeReviews <- readRDS("AnimeReviews/AnimeReviews.RDS")
summary(AnimeReviews)
#head(AnimeReviews)
summ_review <- AnimeReviews %>%
    select(anime_title) %>% 
    group_by(anime_title) %>%
    summarise(Count = n()) %>% 
    arrange(desc(Count))
summary(summ_review$Count)
head(summ_review, 3)
```

```{r plot1}
p1 <- ggplot(AnimeReviews) +
    geom_freqpoly(aes(x=date), bins = 60) +
    xlab("Review Date") + ylab("Number of Reviews") +
    theme_bw()
p1
```

```{r plot2}
review1 <- AnimeReviews[AnimeReviews$anime_title == "Sword Art Online", ]
p2 <- ggplot(review1) +
    geom_freqpoly(aes(x=date), bins = 30) +
    xlab("Review Date") + ylab("Number of Reviews") +
    theme_bw() + ggtitle("Reviews of 'Sword Art Online'") 
p2
```


This anime started airing in Summer 2012, ended in Dec 2012. A short film aired in Dec 2013 and its sequel air in Summer 2014. Rumors of new sequel and movie project started in late 2016.

3. anime info EDA
```{r load data2, warning=F, message=F}
library(visNetwork)
library(car)
AnimeInfo <- readRDS("AnimeInfo/AnimeInfo.RDS")
length(AnimeInfo)
names(AnimeInfo[[1]])

```
```{r plot3_1}
set.seed(0)
graph_list <- sample(AnimeInfo, 100)  ##randomly choose 100 anime for network graph
mklink_lst <- function(lst){
	link_list <- data.frame(from = numeric(), to = numeric(), 
	                        value = numeric(), title = character())
	l <- length(lst)
	#########check common elements in staff and cast members #####################################
	actor_lst <-lapply(lst, function(x) x$anime_mainactors)
	staff_lst <-lapply(lst, function(x) unlist(x$anime_staff))
	for (i in c(1:l-1)){
		for (j in c(i+1:l)){
			temp_actor <- intersect(unlist(actor_lst[i]), unlist(actor_lst[j]))
			temp_staff <- intersect(unlist(staff_lst[i]), unlist(staff_lst[j]))
			if (length(temp_actor) > 0 | length(temp_staff) > 0){
				if (length(temp_actor) > 0){
				    title_actor_temp <- paste0("Cast: ", temp_actor, collapse = '<br>')
				}
				if (length(temp_staff) > 0){
				    title_staff_temp <- paste0("Staff: ", temp_staff, collapse = '<br>')
				}
				if (length(temp_actor) > 0 & length(temp_staff) > 0){
				    title_temp <- paste0(title_staff_temp,'<br>', title_actor_temp)
				    value_t <- length(temp_staff) + length(temp_actor)
				}else if (length(temp_actor) > 0){
				    title_temp <- title_actor_temp
				    value_t <- length(temp_actor)
				}else{
				        title_temp <- title_staff_temp
				        value_t <- length(temp_staff)
				        }
				row <- data.frame(from = i, to = j,
				                  value = value_t, title = title_temp)
			    link_list <- rbind(link_list, row)
			}
		}
	}
	return(link_list)
}

mknode_lst <- function(lst){
    node_list <- data.frame(ID = c(1:length(lst)))
    titles <- unlist(lapply(lst, function(x) x$anime_title))
    node_list$Titles <- titles
    types <- unlist(lapply(lst, function(x) x$anime_type))
    node_list$Types <- types
    members <- unlist(lapply(lst, function(x) x$anime_members))
    members <- as.numeric(gsub("\\,", "", members))
    node_list$Members <- members
    colnames(node_list) <- c("id", "title", "group", "size")
    node_list$size <- 2*sqrt(sqrt(node_list$size))
    node_list$label <- NA
    return(node_list)
}
link_list <- mklink_lst(graph_list)
node_list<- mknode_lst(graph_list)
graph <- visNetwork(nodes = node_list, edges = link_list, main = "Network of 100 Anime") %>%
    visNodes(label = NULL) %>% 
    visLegend(enabled = TRUE, useGroups = TRUE, addNodes = NULL,
              addEdges = NULL, width = 0.2, position = "left", main = NULL) %>% 
    visOptions(selectedBy = "group", highlightNearest = TRUE)
graph
```


4. predict anime score
```{r reconstructing data frame from list}
Title <- unlist(lapply(AnimeInfo, function(x) x$anime_title))
Type <- unlist(lapply(AnimeInfo, function(x) x$anime_type))
Rating <- unlist(lapply(AnimeInfo, function(x) x$anime_rating))
Viewer <-unlist(lapply(AnimeInfo, function(x) as.numeric(gsub("\\,", "", x$anime_members))))
Score <- unlist(lapply(AnimeInfo, function(x) as.numeric(x$anime_score)))
Favorite <- unlist(lapply(AnimeInfo, function(x) as.numeric(gsub("\\,", "", x$anime_favorites))))
Actor1_lst <- lapply(AnimeInfo, function(x) x$anime_mainactor[1])
flag <- unlist(lapply(Actor1_lst, function(x) is.null(x[[1]])))
Actor1_lst[flag] <- NA
Actor1 <- unlist(Actor1_lst)
Actor2_lst <- lapply(AnimeInfo, function(x) x$anime_mainactor[2])
flag <- unlist(lapply(Actor2_lst, function(x) is.null(x[[1]])))
Actor2_lst[flag] <- NA
Actor2 <- unlist(Actor2_lst)
Actor3_lst <- lapply(AnimeInfo, function(x) x$anime_mainactor[3])
flag <- unlist(lapply(Actor3_lst, function(x) is.null(x[[1]])))
Actor3_lst[flag] <- NA
Actor3 <- unlist(Actor3_lst)
Actor4_lst <- lapply(AnimeInfo, function(x) x$anime_mainactor[4])
flag <- unlist(lapply(Actor4_lst, function(x) is.null(x[[1]])))
Actor4_lst[flag] <- NA
Actor4 <- unlist(Actor4_lst)
Director_lst <- lapply(AnimeInfo, function(x) x$anime_staff$Director)
flag <- unlist(lapply(Director_lst, function(x) is.null(x[[1]])))
Director_lst[flag] <- NA
Director <- unlist(Director_lst)
Musician_lst <- lapply(AnimeInfo, function(x) x$anime_staff$Music)
flag <- unlist(lapply(Musician_lst, function(x) is.null(x[[1]])))
Musician_lst[flag] <- NA
Musician <- unlist(Musician_lst)
AnimeInfo_df <- data.frame(Title = Title, Type = Type, Rating = Rating,
                           Director = Director, Actor1 = Actor1, 
                           Actor2 = Actor2, Actor3 = Actor3,
                           Actor4 = Actor4, Musician = Musician,
                           Score = Score,
                           Viewer = Viewer, Favorite = Favorite
                           )
table(AnimeInfo_df$Rating)

#Upon investigation, all animes missing 'Rating' are old animes produced before rating system came and they all look #like "G"-rating animes.
#Set all their ratings to "G - All Ages".
AnimeInfo_df$Rating[AnimeInfo_df$Rating== "None"] <- "G - All Ages"
AnimeInfo_df$Rating <- as.character(AnimeInfo_df$Rating)
AnimeInfo_df$Rating <- as.factor(AnimeInfo_df$Rating)
sum_AnimeInfo <- AnimeInfo_df %>%
    group_by(Type) %>% 
    summarise(MeanScore = mean(Score), SDScore = sd(Score)) %>% 
    arrange(desc(MeanScore))
sum_AnimeInfo
sum_AnimeInfo_Rating <- AnimeInfo_df %>% 
    group_by(Rating) %>% 
    summarise(MeanScore = mean(Score), SDScore = sd(Score)) %>% 
    arrange(desc(MeanScore))
sum_AnimeInfo_Rating
bartlett.test(Score ~ Type, data = AnimeInfo_df) #unequal variance
bartlett.test(Score ~ Rating, data = AnimeInfo_df) #unequal variance
#summary(aov(Score ~ Type, data = AnimeInfo_df))
#summary(aov(Score ~ Rating, data = AnimeInfo_df))
```
```{r plot3_2}
plot1 <- ggplot(AnimeInfo_df) +
    geom_boxplot(aes(y=Score, x= Type, color = Type)) +
    theme_bw() + theme(legend.position="none")
plot1
```
```{r plot3_3}
plot2 <- ggplot(AnimeInfo_df) +
    geom_boxplot(aes(y=Score, x= Rating, color = Rating)) + 
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position="none")
plot2
```
```{r model}
# dummify director variable
# list of anime directors who won awards in last 15-20 years at Tokyo Anime Award or Animation Kobe
direct_award <- c("Daichi, Akitaro", "Miyazaki, Hayao", "Hara, Keiichi", "Kon, Satoshi", 
                  "Miyazaki, Hayao", "Tomino, Yoshiyuki", "Hosoda, Mamoru", "Anno, Hideaki",
                  "Miyazaki, Hayao", "Hosoda, Mamoru", "Yonebayashi, Hiromasa", "Shinbo, Akiyuki",
                  "Hosoda, Mamoru", "Araki, Tetsuro", "Takahata, Isao", "Fujita, YÅichi",
                  "Anno, Hideaki", "Miyazaki, Hayao", "Watanabe, Shinichi", "Daichi, Akitaro",
                  "Okiura, Hiroyuki", "Kitakubo, Hiroyuki", "Hara, Keiichi", "Kuroda, Yosuke",
                  "Kamiyama, Kenji", "Yoshida, Kenichi", "Nagahama, Hiroshi", "Imaishi, Hiroyuki",
                  "Iso, Mitsuo", "Kato, Kunio", "Hosoda, Mamoru", "Okada, Mari",
                  "Agematsu, Noriyasu", "Mizushima, Tsutomu","Kishi, Seiji", "Mizushima, Seiji")
direct_award <- unique(direct_award)
AnimeInfo_df$Award_Director <- ifelse(AnimeInfo_df$Director %in% direct_award, TRUE, FALSE)
## didn't have enough time to do the same thing on actors and musicians
##multi-variable linear regression
model <- glm(Score ~ Award_Director + Viewer + Favorite + Type + Rating, data = AnimeInfo_df)
summary(model)
plot(model)
avPlots(model)
vif(model)
```


Significant coeffiecients: (Intercept), Award_DirectorTRUE, Favorite, Viewer, TypeMusic, TypeONA, TypeOVA, RatingPG-13 - Teens 13 or older, RatingR - 17+ (violence & profanity), RatingR+ - Mild Nudity, RatingRx - Hentai.

Adjusted R-squared:  0.289  only explains 28.9% of the variance in scores of animes.

To do:    
 - Normalize viewer/favorites against time since release;  
 - Use number of favorites on staff/actor pages to turn Director, Actor etc. into continuous variable.
 
 
 
```{r model2, eval=F}
people_df <- read.csv("AnimePeople/AnimePeople.csv", header = T, stringsAsFactors = F)
df<-AnimeInfo_df
df$DirectorScore <- 0
df$ActorScore <- 0
df$MusicianScore <- 0
for (i in c(1:nrow(df))){
    if(!is.na(df$Director[i])){
        df$DirectorScore[i] <- people_df$anime_people_fav[which(people_df$anime_people == df$Director[i])[1]]
    }
    if(!is.na(df$Musician[i])){
        df$MusicianScore[i] <- people_df$anime_people_fav[which(people_df$anime_people == df$Musician[i])[1]]
    }
    if(!is.na(df$Actor1[i])){
        df$ActorScore[i] <- df$ActorScore[i] + people_df$anime_people_fav[which(people_df$anime_people == df$Actor1[i])[1]]
    }
    if(!is.na(df$Actor2[i])){
        df$ActorScore[i] <- df$ActorScore[i] + people_df$anime_people_fav[which(people_df$anime_people == df$Actor2[i])[1]]
    }
    if(!is.na(df$Actor3[i])){
        df$ActorScore[i] <- df$ActorScore[i] + people_df$anime_people_fav[which(people_df$anime_people == df$Actor3[i])[1]]
    }
    if(!is.na(df$Actor4[i])){
        df$ActorScore[i] <- df$ActorScore[i] + people_df$anime_people_fav[which(people_df$anime_people == df$Actor4[i])[1]]
    }
}
#model_lm <- lm(Score ~ ..., data = df)
#after many trials, linear model still doesn't work
#use random forest
```
```{r random forest}
new_df <- readRDS("final_df.RDS")
library(randomForest)
set.seed(0)
train <- sample(c(1:nrow(new_df)), 0.8*nrow(new_df))
rf_df <- randomForest(Score ~ DirectorScore + ActorScore + MusicianScore + Viewer + Favorite + Type + Rating, data = new_df, subset = train, importance = TRUE)
set.seed(0)
oob.err <- numeric(7)
for (mtry in 1:7) {
    fit <- randomForest(Score ~ . - Title , data = new_df[train, ], mtry = mtry)
    oob.err[mtry] <- fit$mse[500]  
    cat("We're performing iteration", mtry, "\n")
}
plot(1:7, oob.err, pch = 16, type = "b",
     xlab = "Variables Considered at Each Split",
     ylab = "OOB Mean Squared Error",
     main = "Random Forest OOB Error Rates\nby # of Variables")
min(oob.err)
which.min(oob.err)
importance(rf_df, type = 1)
varImpPlot(rf_df, type = 1)
rf_df_final <- randomForest(Score ~ DirectorScore +
                                ActorScore + MusicianScore +
                                Viewer + Favorite + Type +
                                Rating, 
                            data = new_df, 
                            subset = train, 
                            mtry = 3)
pred_test <- predict(rf_df_final, new_df[-train,])
Metrics::rmse(new_df$Score[-train], pred_test)
```
