---
title: "Exploring Music Listeners Around the World"
author: "Joshua Litven"
date: "10/14/2016"
output: ioslides_presentation
---
<style type="text/css">
  slides > slide:not(.nobackground):after {
    content: '';
  }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, fig.height=6, fig.align='center')
library(dplyr)
library(ggplot2)
library(lubridate)
library(reshape2)
library(maps)
library(visNetwork)

# Graph settings
TITLE_SIZE = 20
TEXT_SIZE = 20
AXIS_SIZE = 15
WIDTH = 600
HEIGHT = 567
set.seed(2) # recreate results
```

```{r import}
# Import Data -------------------------------------------------------------

# Set the working directory to where the files are stored
setwd('~/Courses/nyc_data_science_academy/bootcamp007_project/Project1-ExploreVis/JoshuaLitven/')
PROCESSED_PATH = file.path('data', 'processed_data')

# Get the artist origins data
artist_locations = readRDS(file.path(PROCESSED_PATH,'artist_locations.rds'))

# Get the user listening history
user_listening_history = readRDS(file.path(PROCESSED_PATH, 'user_listening_history.rds'))
user_listening_history = tbl_df(user_listening_history)
user_listening_history_dems = readRDS(file.path(PROCESSED_PATH, 
                                                'user_listening_history_demographics.rds'))

# Join user listening history and user demographics
user_listening_history_full = left_join(user_listening_history, user_listening_history_dems, by = 'user_id')

# For daily activitiy must filter by areas with utc time zone
utc_countries = c("United Kingdom")
utc_user_listening_history = 
  user_listening_history_full %>%
  filter(country %in% utc_countries)

user_artist_plays = readRDS(file.path(PROCESSED_PATH, 'user_artist_plays.rds'))
user_artist_plays = tbl_df(user_artist_plays)

user_artist_plays_dems = readRDS(file.path(PROCESSED_PATH, 'user_artist_plays_demographics.rds'))
user_artist_plays_dems = tbl_df(user_artist_plays_dems)

# Join artists and demographics
user_artist_plays_full = left_join(user_artist_plays, user_artist_plays_dems, by = "user_id")
```

```{r helper_functions}
# Helper Functions ---------------------------------------------------------

# Find the most popular artists
get_artist_popularity = function(){
  total_num_users = length(unique(user_artist_plays$user_id))
  most_popular_artists = 
    user_artist_plays %>%
    group_by(artist_name) %>%
    summarise(world_pop = n() / total_num_users) %>%
    arrange(desc(world_pop))
  return(most_popular_artists)
}

# Find the most popular artists by each country
get_artist_popularity_by_country = function() {
  
  # Find the number of users in each country
  users_by_country =
    user_artist_plays_full  %>% 
    group_by(country) %>% 
    summarise(num_users = n_distinct(user_id)) %>% 
    arrange(desc(num_users))
  
  # Find the number of listeners for each artist in a country
  artist_pop_by_country =
    user_artist_plays_full %>% 
    group_by(country, artist_name) %>%
    summarise(num_listened = n())
  
  # Find the ratio of users listen to users in each country
  artist_pop_by_country = left_join(artist_pop_by_country, users_by_country, by="country")
  artist_pop_by_country =
    artist_pop_by_country %>% 
    mutate(country_pop = num_listened / num_users)
  return(artist_pop_by_country)
}


# Create popularity choropleth map of a given artist
output_choropleth_map = function(name){
  
  artist_pop_by_country_all = get_artist_popularity_by_country()
  
  artist_pop_by_country =
    artist_pop_by_country_all %>% 
    filter(artist_name == name)
  
  world = map_data("world")
  
  # Fix mappings between user artist plays and world data
  countries = artist_pop_by_country$country
  world_countries = factor(world$region)
  countries = revalue(countries, c("United States"="USA", "Russian Federation"="Russia"))
  artist_pop_by_country$country = countries
  # TODO: Finish fixing
  
  # Join user artist plays with world data
  artist_plays_world = left_join(artist_pop_by_country, world, by=c("country"="region"))
  
  # Cap the country_pop at a maximum popularity level
  max_country_pop = 0.2 
  artist_plays_world$country_pop = pmin(max_country_pop, artist_plays_world$country_pop)
  
  # Plot world with ggplot
  simpleCap <- function(x) {
    s <- strsplit(x, " ")[[1]]
    paste(toupper(substring(s, 1,1)), substring(s, 2),
          sep="", collapse=" ")
  }
  
  p <- ggplot(data = artist_plays_world)
  p <- p + geom_polygon(aes(x=long, 
                            y=lat, 
                            group = group, 
                            fill=country_pop)) +
    theme_minimal() +
    scale_fill_gradient(name="Popularity\n", limits=c(0, max_country_pop), low="#BFEFFF", high="darkblue") +
    theme(plot.title = element_text(size=TITLE_SIZE), text = element_text(size = TEXT_SIZE),
          axis.title.x=element_blank(), axis.ticks=element_blank(), axis.text=element_blank()) +
    labs(title=paste("Users Listening to",simpleCap(name)), x="", y="")
  p
}

# Compare world popularity and country popularity for selected artists
create_popularity_plot = function(artist_popularity){

  country_name = as.character(artist_popularity$country[1])
  artist_popularity_long = melt(artist_popularity, id.vars=c("country", "artist_name"))
  artist_popularity_long$artist_name = factor(artist_popularity_long$artist_name,
                                              levels=artist_popularity$artist_name)
  
  # Plot world and origin popularity of artists
  g = ggplot(artist_popularity_long) +
    geom_bar(aes(x = artist_name, y = value, fill = variable), 
             stat="identity", position = "dodge") +
    theme_minimal() +
    theme(plot.title = element_text(size=TITLE_SIZE), text = element_text(size = TEXT_SIZE),
          axis.text.x=element_text(size=AXIS_SIZE, angle = 90, hjust = 1)) +
    scale_fill_discrete(name="", labels=c(country_name, "World")) +
    labs(x = "", y = "Popularity\n", title = paste("Popularity of Top Artists in ", country_name))
  return(g)
}
```

## Introduction

* As the world becomes increasingly globalized, it is crucial to understand other cultures as much as possible.

* In NYC, it is clear we do that by trying the world's finest cuisine.

* But what about connecting through music?

## Goal

*Find artists popular in different countries but relatively unknown to the rest of the world.*

## Data

Data on last.fm users were scraped to create two data sets:

* Data Set 1: User listening history for 1000 users
* Data Set 2: User favorite artist plays for 360k users

## Question One

*How do users listen to artists?*

##
```{r}
# What is a Random user's weekly listening activity? -----------------------------

# Pick a random user
sampled_user_id = sample(utc_user_listening_history$user_id, 1)

sampled_user_listening_history =
  utc_user_listening_history %>%
  filter(user_id == sampled_user_id)

# Pick a random date
start = date(sample(sampled_user_listening_history$date, 1))
end = start + days(6)

week_listening_history = 
  sampled_user_listening_history %>%
  filter(date >= start & date <= end)

# View their listening amount summed across weekdays
p = ggplot(data=week_listening_history, aes(x=hour)) +
  geom_bar(aes(fill=DofW)) +
  facet_wrap( ~ DofW, ncol = 1) +
  labs(title="Random User's Listening Activity (1 Week)", x="\nHour", y="Num Songs\n")  +
  scale_y_continuous(breaks=c(0,15)) +
  scale_fill_discrete(name="") +
  theme_minimal() +
  theme(plot.title = element_text(size=TITLE_SIZE), 
        text = element_text(size = TEXT_SIZE),
        axis.text=element_text(size=10), strip.text.x = element_blank())
p
```

##
```{r}
# What is the user's trend for weekly listening activity? ------------------------------------------
png(file="mygraphic.png",width=WIDTH,height=HEIGHT)
p = ggplot(data=subset(sampled_user_listening_history, !is.na(hour)), aes(x=hour)) +
  geom_bar(aes(fill=DofW)) +
  facet_wrap( ~ DofW, ncol = 1) +
  labs(title="Random User's Listening Activity (Aggregated)\n", x="\nHour", y="Num Songs\n") +
  scale_fill_discrete(name="") +
  theme_minimal() +
  theme(plot.title = element_text(size=TITLE_SIZE), text = element_text(size = TEXT_SIZE),
        axis.text=element_text(size=10), strip.text.x = element_blank())
p
dev.off()
```

##
```{r}
# Break down by most common artists
png(file="plots/random_user_listening_activity_breakdown.png",width=WIDTH,height=HEIGHT)
sampled_user_artists = 
  sampled_user_listening_history %>% 
  group_by(artist_name) %>% 
  tally(sort = TRUE)

# Get top percentile
ratio = .1
percent = ratio * 100
num_artists = as.integer(ratio * nrow(sampled_user_artists))
most_common_artists =
  sampled_user_artists %>% 
  top_n(num_artists, n)
most_common_artists = as.character(most_common_artists$artist_name)

sampled_user_listening_history_top =
  sampled_user_listening_history %>%
  mutate(top = artist_name %in% most_common_artists)

p = ggplot(data=subset(sampled_user_listening_history_top, !is.na(hour)), aes(x=hour)) +
  geom_bar(aes(fill=top)) +
  facet_wrap( ~ DofW, ncol=1) +
  labs(title="Random User's Listening Activity (Aggregated)\n", x="\nHour", y="Num Songs\n") +
  scale_y_continuous(breaks=c(0,500)) +
  scale_fill_discrete(name="Artists", labels=c(paste0("Bottom ", 100 - percent, "%"),
                                               paste0("Top ", percent, "%"))) +
  theme_minimal() +
  theme(plot.title = element_text(size=TITLE_SIZE), text = element_text(size = TEXT_SIZE),
        axis.text=element_text(size=10), strip.text.x = element_blank())
p
dev.off()
```

##
```{r}
# How diversified are users? ----------------------------------------------

# Calculate the cumulative sum of all artists for every user
user_artist_counts =
  user_listening_history %>%
  group_by(user_id, artist_name) %>% 
  summarise(count = n())

# Plot the user's cumulative distributions
png(file="plots/diversity_of_user_history.png",width=WIDTH,height=HEIGHT)
g = ggplot(user_artist_counts, aes(count)) +
  stat_ecdf(geom = "step", aes(fill=user_id), alpha=.5) +
  coord_cartesian(xlim=c(50,1000)) +
  labs(title="Diversity of User History", x="\nNum Artists", y="% of History\n") +
  theme_minimal() +
  theme(plot.title = element_text(size=TITLE_SIZE), text = element_text(size = TEXT_SIZE))
g

# Output plot
dev.off()
```

## 
```{r}
# Compare the median and most diverse user to the number of artists

user_distinct_artist_counts = 
  user_listening_history %>%
  group_by(user_id) %>%
  summarise(distinct_artists = n_distinct(artist_name))

median_diverse_user_artist_count = median(user_distinct_artist_counts$distinct_artists)
most_diverse_user_artist_count =  max(user_distinct_artist_counts$distinct_artists)
num_unique_artists = length(unique(user_listening_history$artist_id))

artist_counts = data.frame(count=c(median_diverse_user_artist_count, 
                                   most_diverse_user_artist_count, 
                                   num_unique_artists),
                           name=c("Median Diversity", "Max Diversity", "Total"))
g = ggplot(artist_counts) +
  geom_bar(aes(x=reorder(name, count), y=count, fill=name), 
           stat="identity") +
  labs(title="User Diversity vs. Number of Artists", x="", y="Num Artists\n") +
  theme_minimal() +
  theme(plot.title = element_text(size=TITLE_SIZE), text = element_text(size = TEXT_SIZE),
        axis.text=element_text(size=AXIS_SIZE)) +
  scale_fill_discrete(guide = FALSE)
g
```

## Question Two

*Where do users listen to artists?*

## 
```{r}
# Who are the most popular artists? ---------------------------------------

num_artists = 25
most_popular_artists = 
  get_artist_popularity() %>% 
  slice(1:num_artists)

# Plot the most popular artists
png(file="plots/most_popular_artists.png",width=WIDTH,height=HEIGHT)
g = ggplot(data=most_popular_artists, aes(x = artist_name)) +
  geom_bar(aes(x = reorder(artist_name, world_pop), y = world_pop, fill=artist_name), 
           stat = "identity") +
  coord_flip() +
  labs(title="Most Popular Artists", x="Artist\n", y="\nWorldwide Popularity") +
  theme_minimal() +
  scale_fill_discrete(guide=FALSE) +
  theme(plot.title = element_text(size=TITLE_SIZE), text = element_text(size = TEXT_SIZE))
g
dev.off()
```

```{r}
# Where do users come from?

# Get users by country
n_users = nrow(user_artist_plays_dems)
users_by_country =
  user_artist_plays_dems %>% 
  group_by(country) %>% 
  tally() %>% 
  mutate(ratio = n / n_users) %>% 
  arrange(desc(ratio))

world = map_data("world")

# Fix mappings between user artist plays and world data
countries = users_by_country$country
world_countries = factor(world$region)
countries = plyr::revalue(countries, c("United States"="USA", 
                                       "Russian Federation"="Russia",
                                       "United Kingdom"="UK",
                                       "Korea, Republic of"="South Korea",
                                       "Viet Nam"="Vietnam",
                                       "Lao People's Democratic Republic"="Laos",
                                       "Palestinian Territory, Occupied"="Palestine",
                                       "Congo, the Democratic Republic of the"="Democratic Republic of the Congo",
                                       "Iran, Islamic Republic of"="Iran",
                                       "Macao"="China",
                                       "Syrian Arab Republic"="Syria",
                                       "Congo"="Democratic Republic of the Congo",
                                       "Hong Kong"="China",
                                       "Tanzania, United Republic of"="Tanzania",
                                       "Falkland Islands (Malvinas)"="Falkland Islands",
                                       "Libyan Arab Jamahiriya"="Libya"))
users_by_country$country = countries

# Join user artist plays with world data
users_by_country_world = left_join(world, users_by_country, by=c("region"="country"))

png(file="plots/users_choropleth.png",width=800,height=531)
p <- ggplot(data = users_by_country_world)
p <- p + geom_polygon(aes(x=long, 
                          y=lat, 
                          group = group, 
                          fill=ratio)) +
  theme_minimal() +
  scale_fill_gradient(name="Proportion of Users\n", low="#BFEFFF", high="darkblue", na.value="white") +
  theme(plot.title = element_text(size=TITLE_SIZE), text = element_text(size = TEXT_SIZE),
        axis.title.x=element_blank(), axis.ticks=element_blank(), axis.text=element_blank()) +
  labs(title="User Demographics", x="", y="") 
p
dev.off()
```

## 
```{r}
# Where is most popular artist listened to? -------------------------------

top_artist = get_artist_popularity()$artist_name[1]
output_choropleth_map(top_artist)
```

## 
```{r}
# Where is a random artist listened to? -----------------------------------

random_artist = "boards of canada"
output_choropleth_map(random_artist)
```

##
```{r}
# How does popularity in origin relate to world popularity? ---------------

# Join with country popularity
artist_popularity_by_country = get_artist_popularity_by_country()

artist_popularity_locations = left_join(artist_locations, artist_popularity_by_country,
                                        by= c("artist_name"="artist_name",
                                              "country"="country"))
artist_popularity_locations =
  artist_popularity_locations %>% 
  select(country, artist_name, country_pop) %>% 
  na.omit()

# Join with world popularity
artist_world_popularity = get_artist_popularity()

artist_popularity_full =left_join(artist_popularity_locations, artist_world_popularity, 
                                  by="artist_name")

# Plot country popularity vs world popularity
ggplot(artist_popularity_full, aes(y=country_pop, x=pop)) +
  geom_point() +
  coord_cartesian(xlim=c(0,0.2), ylim=c(0,0.2)) +
  stat_function(fun=function(x) x, geom = "line") +
  theme_minimal() +
  theme(plot.title = element_text(size=TITLE_SIZE), text = element_text(size = TEXT_SIZE),
        axis.text=element_text(size=AXIS_SIZE)) +
  labs(title="Origin Popularity vs. World Popularity", 
       x="\nWorld Popularity", y="Origin Popularity\n")
```

##
```{r}
# How do popular artists from different countries compare? ----------------

# Get most popular artists from a selection of countries
artist_popularity_by_country = get_artist_popularity_by_country()
countries = c("Canada", "France", "China", "United States", "Brazil")
num_artists = 5

artist_popularity_selected_countries = 
  artist_popularity_by_country %>% 
  filter(country %in% countries) %>% 
  select(country, artist_name, country_pop) %>% 
  top_n(num_artists, country_pop)

# Find the repeated artists
repeated_artists =
  artist_popularity_selected_countries %>% 
  group_by(artist_name) %>% 
  filter(n() > 1) %>%
  select(artist_name)
repeated_artists = as.vector(repeated_artists[[1]])

# Create factor indicating whether the artist is repeated
artist_popularity_selected_countries_long = melt(artist_popularity_selected_countries, 
                                                 id.vars=c("country", "artist_name"))

artist_popularity_selected_countries_full =
  artist_popularity_selected_countries_long %>%
  mutate(repeated = artist_name %in% repeated_artists) %>% 
  arrange((desc(value)))

# Plot top artists across selected countries
png(file="plots/top_artists_across_countries.png",width=WIDTH,height=HEIGHT)
ggplot(data = artist_popularity_selected_countries_full, aes(x=artist_name, y=value)) +
  geom_bar(stat="identity", aes(fill=artist_name)) +
  facet_wrap( ~ country, nrow=length(countries)) +
  theme_minimal() +
  scale_fill_discrete(name="") +
  theme(plot.title = element_text(size=TITLE_SIZE), text = element_text(size = TEXT_SIZE),
        axis.text.y=element_text(size=10),axis.text.x=element_blank()) +
  labs(x = "", y = "Popularity\n", title = "Top Artists Across Countries")
dev.off()
```


##
```{r similarity of users in same country vs foreign}
# How similar are users in the same country compared to foreign users?

similarity_metric = function(user1, user2){
  user1_artists = user_artist_plays %>% filter(user_id == user1) %>% .$artist_name
  user2_artists = user_artist_plays %>% filter(user_id == user2) %>% .$artist_name
  
  common_artists = intersect(user1_artists, user2_artists)
  
  similarity_metric = length(common_artists) / (length(user1_artists) * length(user2_artists))
}

meanSimilarity = 0
for(i in 1:1000){
  # Sample a user
  sampled_user = sample_n(user_artist_plays_full, 1)
  
  # Sample another user from the same country
  other_user = sample_n(user_artist_plays_full %>% filter(country == sampled_user$country), 1)
  
  meanSimilarity = meanSimilarity + similarity_metric(sampled_user$user_id, other_user$user_id)
}

```
## Question Three

*How to find artists popular in a country unknown to the rest of the world?*

```{r}
# Start with the popularity of artists for each country
artist_popularity_by_country = get_artist_popularity_by_country()

# Join with world popularity
artist_world_popularity = get_artist_popularity()

artist_popularity_full = left_join(artist_popularity_by_country, artist_world_popularity, 
                                  by="artist_name")
```

##
```{r}
# Plot example country
example_country = "France"
num_artists = 10

artist_popularity_example =
  artist_popularity_full %>% 
  filter(country==example_country) %>% 
  arrange(country_pop) %>% 
  top_n(num_artists, country_pop) %>%
  select(country, artist_name, country_pop, world_pop) %>%
  arrange(desc(country_pop))

g = create_popularity_plot(artist_popularity_example)
g
```

##
```{r}
# Let's clip artists with high world pop
world_pop_cut_off = 0.1

# Show world popularity cut off 
g + geom_hline(yintercept = world_pop_cut_off) + 
  annotate("text", x=num_artists - 1, y=0.16, label="Popularity Cutoff") +
  annotate("segment", x=num_artists - 1, xend=num_artists - 1, y=0.15, 
           yend=0.1, colour="black", size=1, arrow=arrow()) 

```

##
```{r}
artist_popularity_example_clipped =
  artist_popularity_full %>% 
  filter(country==example_country, world_pop < world_pop_cut_off) %>% 
  arrange(country_pop) %>% 
  top_n(num_artists, country_pop) %>%
  select(country, artist_name, country_pop, world_pop) %>%
  arrange(desc(world_pop))

create_popularity_plot(artist_popularity_example_clipped)
```

## How to Find Unknown Aritsts {.flexbox .vcenter}
*net popularity = country popularity - world popularity*

##
```{r}
# Let's sort by net_pop = country_pop - world_pop
artist_popularity_example_net = 
  artist_popularity_full %>% 
  filter(country==example_country, world_pop < world_pop_cut_off) %>% 
  mutate(net_pop = country_pop - world_pop) %>%
  arrange(desc(net_pop)) %>% 
  top_n(num_artists, net_pop) %>%
  select(country, artist_name, country_pop, world_pop)
  
create_popularity_plot(artist_popularity_example_net) +
  labs(title="Popularity of Top Artists in France (Reordered)")
```

##
```{r}
# Get top net pop artists for each country
artist_popularity_net = 
  artist_popularity_full %>% 
  filter(world_pop < world_pop_cut_off) %>%
  mutate(net_pop = country_pop - world_pop) %>% 
  arrange(desc(net_pop)) %>% 
  top_n(num_artists, net_pop) %>% 
  select(country, artist_name, country_pop, world_pop)

# Canada
create_popularity_plot(artist_popularity_net %>% filter(country=="Canada"))
```

## 
```{r}
# China
create_popularity_plot(artist_popularity_net %>% filter(country=="China"))
```

##
```{r}
# USA
create_popularity_plot(artist_popularity_net %>% filter(country=="United States"))
```

```{r create_youtube_playlist}
# Load the playlist creator
setwd('~/Courses/nyc_data_science_academy/bootcamp007_project/Project1-ExploreVis/JoshuaLitven/')
source('code/create_youtube_playlist.R')

# Start with the popularity of artists for each country
artist_popularity_by_country = get_artist_popularity_by_country()

# Get niche artists by country
# Join with world popularity
artist_world_popularity = get_artist_popularity()

artist_popularity_full = left_join(artist_popularity_by_country, artist_world_popularity, 
                                  by="artist_name")
num_users_threshold = 10
num_artists = 10
artist_popularity_net = 
  artist_popularity_full %>% 
  filter(num_users > num_users_threshold) %>% 
  mutate(net_pop = country_pop - world_pop) %>% 
  arrange(desc(net_pop)) %>% 
  top_n(num_artists, net_pop) %>% 
  select(country, artist_name, country_pop, world_pop, net_pop)

num_niche_artists = 1
niche_artists_by_country = 
  artist_popularity_net %>%
  group_by(artist_name) %>% 
  filter(n() == 1) %>%
  ungroup(artist_name) %>% 
  group_by(country) %>% 
  top_n(num_niche_artists, net_pop) %>% 
  arrange(country)

create_playlist(niche_artists_by_country)
```

```{r create artist-country network graph}

# Create the artist country network graph from a dataframe
# The dataframe contains columns: country, artist_name, country_pop
# If use_pop=TRUE, artist node size is proportional to country_pop
create_artists_country_network = function(artists_by_country, title, use_pop=TRUE, show_labels=TRUE,
                                          select_option=TRUE){
  # Edges are defined from country to artist
  edges = data.frame(from=artists_by_country$country,
                     to=artists_by_country$artist_name)
  
  # Get the counts of each artist to specify size in network
  artist_counts = artists_by_country  %>% group_by(artist_name)  %>% count(artist_name)
  artist_pop = artists_by_country %>% ungroup(artist_name) %>% .$country_pop
  country_counts = artists_by_country %>%  group_by(country) %>% count(country)
  country_counts$country = as.character(country_counts$country)
  
  # Specify the nodges
  # The nodes are all countries and artists
  nodes = data.frame(id = c(artist_counts$artist_name, 
                            country_counts$country),
                     group = c(rep("Artist", nrow(artist_counts)),
                               rep("Country", nrow(country_counts))))
  nodes$shape <- "dot"  
  nodes$shadow <- TRUE # Nodes will drop shadow
  nodes$title <- c(artist_counts$artist_name, country_counts$country)
  if(!show_labels){
    nodes$label = NA
  }
  nodes$borderWidth <- 2 # Node border width
  nodes$value = ifelse(rep(use_pop, nrow(nodes)), c((10 * artist_pop)^2, rep(1, nrow(country_counts))), 
                       c(artist_counts$n, rep(1, nrow(country_counts))))
  
  nodes$color.border <- "black"
  nodes$color.highlight.background <- "orange"
  nodes$color.highlight.border <- "darkred"
  
  network = 
    visNetwork(nodes, edges, main=title) %>%
    visOptions(highlightNearest = TRUE) %>% 
    visGroups(groupname = "Country", color = "red") %>%
    visGroups(groupname = "Artist", color = "lightblue") %>% 
    visLegend()
  if(select_option){
    network =
      network %>% 
      visOptions(highlightNearest = TRUE,
                 nodesIdSelection = list(values=country_counts$country))
  }
  return(network)
}

# Start with the popularity of artists for each country
artist_popularity_by_country = get_artist_popularity_by_country()

# Get top artists by country
num_artists = 10
top_artists_by_country = 
  artist_popularity_by_country %>%
  group_by(country) %>% 
  top_n(num_artists, country_pop)

# Sample for cleaner network plot
sample_user_threshold = 500
sampled_artists_by_country = 
  top_artists_by_country %>% 
  filter(num_users > sample_user_threshold) %>% 
  select(country, artist_name, country_pop)

# Join with world popularity to get niche artists
artist_world_popularity = get_artist_popularity()
artist_popularity_full = left_join(artist_popularity_by_country, artist_world_popularity, 
                                  by="artist_name")
num_users_threshold = 10
artist_popularity_net = 
  artist_popularity_full %>% 
  filter(num_users > num_users_threshold) %>% 
  mutate(net_pop = country_pop - world_pop) %>% 
  arrange(desc(net_pop)) %>% 
  top_n(num_artists, net_pop) %>% 
  select(country, artist_name, country_pop, world_pop, net_pop)

num_niche_artists = 10
niche_artists_by_country = 
  artist_popularity_net %>%
  group_by(artist_name) %>% 
  filter(n() == 1) %>%
  ungroup(artist_name) %>% 
  group_by(country) %>% 
  top_n(num_niche_artists, net_pop)

sampled_niche_artists_by_country = 
  niche_artists_by_country %>% 
  filter(country %in% sampled_artists_by_country$country)

# Have to manually save image
network = create_artists_country_network(niche_artists_by_country, 
                                         use_pop=TRUE, show_labels=TRUE,
                                         select_option=TRUE,
                                         title="Niche Artist Network")
network

# Save to file
network %>% visSave(file = "niche_artist_network.html")
```

## Coming Soon
A youtube playlist with top artists from each country!

## Future Work
* Look up when artists started to discover new bands
* Investigate why artists become popular in certain countries
* Ineractive map to play music popular in a country
    + (Shiny Project???)
  
## References
* Last.fm Data
    + [Music Recommendation Datasets for Research](http://www.dtic.upf.edu/~ocelma/MusicRecommendationDataset/lastfm-1K.html)
* Artist Location Data
    + [Million Song Dataset](http://labrosa.ee.columbia.edu/millionsong/)