---
title: "An Analysis of NYC Public School SAT scores"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(openxlsx)
library(dplyr)
library(ggplot2)
library(GGally)
library(grid)

school_quality5 <- read.xlsx("2014_2015_HS_SQR_Results_2016_01_07.xlsx", sheet = 5, startRow = 2, colNames = TRUE)
SAT_2015 = 
  school_quality5 %>%
  select(
    DBN,
    School.Name,
    Avg.Math = Average.Score.SAT.Math,
    Avg.Reading = Average.Score.SAT.Critical.Reading,
    Avg.Writing = Average.Score.SAT.Writing
  ) %>%
  mutate (
    Avg.Total = Avg.Math + Avg.Reading + Avg.Writing  # add a new column for average total SAT score
  )

school_quality1 <- read.xlsx("2014_2015_HS_SQR_Results_2016_01_07.xlsx", sheet = 1, startRow = 2, colNames = TRUE)
summary_2015 = school_quality1 %>%
  select(1,2,4, 12:17, 26:40)

SAT_summary_2015 = inner_join(SAT_2015, summary_2015, by = "DBN")


```

## Introduction

This project will explore data that New York City publishes on its public school system. NYC provides a very large amount of open data on its website, including 245 datasets that are categorized within the [education](https://nycopendata.socrata.com/data?cat=education) section. This project will focus on several data sets:

* [2014-2015 School Quality Reports Results For High Schools](https://data.cityofnewyork.us/Education/2014-2015-School-Quality-Reports-Results-For-High-/vrfr-9k4d) (includes SAT scores for each school and extensive data on the learning environment)

* [DOE High School Directory 2014-2015](https://data.cityofnewyork.us/Education/DOE-High-School-Directory-2014-2015/n3p6-zve2) (includes location data of of schools)



## SAT Data

First, let's see how the SAT data looks:


```{r, echo = F, warning = F, message = F}

ggplot(SAT_summary_2015, aes(x=Avg.Total)) + 
  geom_histogram(binwidth = 100, 
                 col="white", 
                 fill="steelblue") + 
  labs(title="NYC Public Schools SAT Score Histogram") +
  labs(x="Average Score for School", y="Count") +
  geom_vline(aes(xintercept=mean(Avg.Total, na.rm=T)),   # Ignore NA values for mean
             color="darkgreen", linetype="dashed", size=1) +
  geom_text(
    mapping = aes(x2, y2, label = text2),
    data = data.frame(x2 = 1290, y2 = 110, text2 = "mean[score]==1273"),
    color = I("darkgreen"),
    parse = TRUE,
    hjust = 0
  ) +
  geom_vline(aes(xintercept=median(Avg.Total, na.rm=T)),   # Ignore NA values for mean
             color="purple", linetype="dashed", size=1) +
  geom_text(
    mapping = aes(x2, y2, label = text3),
    data = data.frame(x2 = 1244, y2 = 120, text3 = "median[score]==1227"),
    color = I("purple"),
    parse = TRUE,
    hjust = 0
  )


#summary statistics
value = c(
  min(SAT_summary_2015$Avg.Total, na.rm = T),
  max(SAT_summary_2015$Avg.Total, na.rm = T),
  mean(SAT_summary_2015$Avg.Total, na.rm = T),
  sum(SAT_summary_2015$Avg.Total*SAT_summary_2015$Enrollment, na.rm=T)/sum(SAT_summary_2015$Enrollment),
  median(SAT_summary_2015$Avg.Total, na.rm = T),
  sd(SAT_summary_2015$Avg.Total, na.rm = T)
)
function. = c("min", "max", "mean", "weighted average", "median", "std dev")
summary_stats = data.frame(function., value)
summary_stats


```

In total, there are 491 public schools in the dataset. Of the school, 82 (or 16.7%) are missing data for average SAT score. Each SAT score is comprised of results from three section: Math, Reading, and Writing. The scores for these subsections have similar distributions, as seen in the violin plots below: 

```{r, echo = F, warning = F, message = F}

section_scores = select(SAT_summary_2015, Avg.Math, Avg.Reading, Avg.Writing)
section_scores2 <- reshape2::melt(section_scores, id.vars = NULL)
section_scores2 <- filter(section_scores2, value != "NA")

#calculate the means of each section to display on violin plot
means <- aggregate(value ~  variable, section_scores2, mean)
means[,2] = round(means$value)

#plot violin plot with means displayed
ggplot(section_scores2, aes(x = variable, y = value, fill = variable)) + 
  geom_violin() +
  geom_boxplot(width = 0.3) +
  stat_summary(fun.y=mean, colour="darkblue", geom="point", 
               shape=18, size=3, show.legend = FALSE) +
  geom_text(data = means, aes(label = value, y = value + 12))

```

The boxplots show that the average for the Math section is the highest (432), followed by Reading (423), and Writing (418). Math also has several higher outliers, indicating that a few schools have relative strength in Math.

If we run a correlation, we can see that the three scores are highly correlated with each other so a school that has a high average score on one section will very likely have a similar score on the other sections. As we might expect, the Reading and Writing sections are more correlated with each other than with the Math section. Reading and Writing test similar skills, so a school that does well on one could be expected to do well on the other.

```{r, echo = F, warning = F, message = F}

cor(SAT_summary_2015[complete.cases(SAT_summary_2015), c("Avg.Math", "Avg.Reading", "Avg.Writing")],)

#pair plot
ggpairs(
  SAT_summary_2015[complete.cases(SAT_summary_2015), c("Avg.Math", "Avg.Reading", "Avg.Writing")],
  upper = list(continuous = wrap("cor", size = 10, color = "steelblue")), 
  lower = list(continuous = wrap("smooth", color = "steelblue")),
  diag = list(continuous = wrap("barDiag", bins = 100, color = "steelblue")),
  axisLabels = 'show') +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    axis.ticks = element_blank())

```


## Top and Bottom Schools

Let's take a closer look at the schools that have SAT score in the top ten 10% for NYC:

```{r, echo = F, warning = F, message = F}

top =
  SAT_summary_2015 %>%
  select(School.Name.x, Avg.Total, Avg.Math, Avg.Reading, Avg.Writing) %>%
  arrange(desc(Avg.Total)) %>%
  top_n(37)


ggplot(top, aes(x = reorder(School.Name.x, Avg.Total), y = Avg.Total)) +
  geom_bar(stat = "identity", col = "white", fill = "steelblue") +
  coord_flip(ylim=c(1400, 2200)) +
  labs(title = "Top 10% NYC Public Schools SAT Scores", x = "", y = "Average SAT Score for School")

ggplot(top, aes(x = reorder(School.Name.x, Avg.Total), y = Avg.Math)) +
  geom_bar(stat = "identity", col = "white", fill = "steelblue") +
  coord_flip(ylim=c(475, 800)) +
  labs(title = "Top 10% NYC Public Schools SAT Scores", x = "", y = "Average Math SAT Score for School")

ggplot(top, aes(x = reorder(School.Name.x, Avg.Total), y = Avg.Reading)) +
  geom_bar(stat = "identity", col = "white", fill = "steelblue") +
  coord_flip(ylim=c(475, 800)) +
  labs(title = "Top 10% NYC Public Schools SAT Scores", x = "", y = "Average Reading SAT Score for School")

ggplot(top, aes(x = reorder(School.Name.x, Avg.Total), y = Avg.Writing)) +
  geom_bar(stat = "identity", col = "white", fill = "steelblue") +
  coord_flip(ylim=c(475, 800)) +
  labs(title = "Top 10% NYC Public Schools SAT Scores", x = "", y = "Average Writing SAT Score for School")
```

Similarly, we can look at schools in the bottom 10%. Note that there is a small but significant number of schools that have "NA" for the SAT score (16.7%), so this graphic only shows the schools that have an SAT score listed:

```{r, echo = F, warning = F, message = F}

bottom =
  SAT_summary_2015 %>%
  select(School.Name.x, Avg.Total, Avg.Math, Avg.Reading, Avg.Writing) %>%
  arrange(Avg.Total) %>%
  filter(Avg.Total != "NA") %>%
  top_n(-37)

ggplot(bottom, aes(x = reorder(School.Name.x, -Avg.Total), y = Avg.Total)) +
  geom_bar(stat = "identity", col = "white", fill = "steelblue") +
  coord_flip(ylim=c(900, 1200)) +
  labs(title = "Bottom 10% NYC Public Schools SAT Scores", x = "", y = "Average SAT Score for School")

ggplot(bottom, aes(x = reorder(School.Name.x, -Avg.Total), y = Avg.Math)) +
  geom_bar(stat = "identity", col = "white", fill = "steelblue") +
  coord_flip(ylim=c(275, 500)) +
  labs(title = "Bottom 10% NYC Public Schools SAT Scores", x = "", y = "Average Math SAT Score for School")

ggplot(bottom, aes(x = reorder(School.Name.x, -Avg.Total), y = Avg.Reading)) +
  geom_bar(stat = "identity", col = "white", fill = "steelblue") +
  coord_flip(ylim=c(275, 500)) +
  labs(title = "Bottom 10% NYC Public Schools SAT Scores", x = "", y = "Average Reading SAT Score for School")

ggplot(bottom, aes(x = reorder(School.Name.x, -Avg.Total), y = Avg.Writing)) +
  geom_bar(stat = "identity", col = "white", fill = "steelblue") +
  coord_flip(ylim=c(275, 500)) +
  labs(title = "Bottom 10% NYC Public Schools SAT Scores", x = "", y = "Average Writing SAT Score for School")


```

What causes schools to be in the top 10% or bottom 10%? Are there similarities between schools that score similarly?

##School Characteristics

The dataset also includes many data points on different characteristics of each school. Let's take a look at some of these data.

```{r, echo = F, warning = F, message = F}

colnames(SAT_summary_2015)[8:29]

```


One charachteristic that I thought might be important was a value called "Rigorous.Instruction.Percent.Positive".

```{r, echo = F, warning = F, message = F}

#first adjust name of column to call
names(SAT_summary_2015)[names(SAT_summary_2015) == "Rigorous.Instruction.-.Percent.Positive"] <- 'Rigorous.Instruction.Percent.Positive' 

ggplot(SAT_summary_2015, aes(x = Rigorous.Instruction.Percent.Positive, y = Avg.Total)) +
  geom_point(color="steelblue") +
  geom_smooth(method = "lm", color = "darkgreen")

```

Although there is a slightly correlation, it is not nearly as correlated as student attendance rate.

```{r, echo = F, warning = F, message = F}

ggplot(SAT_summary_2015, aes(x = Student.Attendance.Rate, y = Avg.Total)) +
  geom_point(color="steelblue") +
  geom_smooth(method = "lm", color = "darkgreen")

```

A strongly negatively correlated characteristic is the economic need index of a school's students.

```{r, echo = F, warning = F, message = F}

ggplot(SAT_summary_2015, aes(x = Economic.Need.Index, y = Avg.Total)) +
  geom_point(color="steelblue") +
  geom_smooth(method = "lm", color = "darkgreen")

```

Below are the relative correlations of all the characteristics with a school's average SAT score.


```{r, echo = F, warning = F, message = F}

correlation_tot = sapply(SAT_summary_2015[complete.cases(SAT_summary_2015), -c(1:7)], function(x)
  cor(x, SAT_summary_2015[complete.cases(SAT_summary_2015), "Avg.Total"]))
correlation_tot = data.frame(labels(correlation_tot), unname(correlation_tot))
colnames(correlation_tot) = c("factor", "Avg.Total")

correlation_math = sapply(SAT_summary_2015[complete.cases(SAT_summary_2015), -c(1:7)], function(x)
  cor(x, SAT_summary_2015[complete.cases(SAT_summary_2015), "Avg.Math"]))
correlation_math = data.frame(labels(correlation_math), unname(correlation_math))
colnames(correlation_math) = c("factor", "Avg.Math")

correlation_reading = sapply(SAT_summary_2015[complete.cases(SAT_summary_2015), -c(1:7)], function(x)
  cor(x, SAT_summary_2015[complete.cases(SAT_summary_2015), "Avg.Reading"]))
correlation_reading = data.frame(labels(correlation_reading), unname(correlation_reading))
colnames(correlation_reading) = c("factor", "Avg.Reading")

correlation_writing = sapply(SAT_summary_2015[complete.cases(SAT_summary_2015), -c(1:7)], function(x)
  cor(x, SAT_summary_2015[complete.cases(SAT_summary_2015), "Avg.Writing"]))
correlation_writing = data.frame(labels(correlation_writing), unname(correlation_writing))
colnames(correlation_writing) = c("factor", "Avg.Writing")

correlationdf = data.frame(factor = correlation_tot$factor,
                      Avg.Total = correlation_tot$Avg.Total,
                      Avg.Math = correlation_math$Avg.Math,
                      Avg.Reading = correlation_reading$Avg.Reading,
                      Avg.Writing = correlation_writing$Avg.Writing) 


correlationdfsorted <- reshape2::melt(correlationdf[,c("factor","Avg.Total","Avg.Math", "Avg.Reading", "Avg.Writing")],id.vars = 1)


ggplot(correlationdf, aes(reorder(factor, Avg.Total), Avg.Total)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Correlation of School Characteristics To School's Average SAT Score",
       x = "", y = "Correlation") +
  coord_flip(ylim = c(-0.75, 0.75)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
  
ggplot(correlationdfsorted, aes(reorder(factor, value), value)) +
  geom_bar(stat = "identity", aes(fill = variable), position = "dodge") +
  labs(title = "Correlation of School Characteristics To School's Average SAT Score",
       x = "", y = "Correlation") +
  coord_flip(ylim = c(-0.75, 0.75)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

We see that the highest positively correlated characteristic is a school's attendance rate, while the top negatively correlated characteristic is 
* the economic need index of a school's students (look up)
* % HRA eligibles
* % of chronic absences (more than a certain amount of days)

Essentially, student? poverty is the most negatively correlared charachteristic, while student attendance is the most positively correlated characteristic. 

If we look at the top 

## Race

```{r, echo = F, warning = F, message = F}

correlation_race = sapply(SAT_summary_2015[complete.cases(SAT_summary_2015), c(21:24)], function(x)
  cor(x, SAT_summary_2015[complete.cases(SAT_summary_2015), "Avg.Total"]))
correlation_race = data.frame(labels(correlation_race), unname(correlation_race))
colnames(correlation_race) = c("race", "Avg.Total")
correlation_race



correlationdf = data.frame(factor = correlation_tot$factor,
                      Avg.Total = correlation_tot$Avg.Total,
                      Avg.Math = correlation_math$Avg.Math,
                      Avg.Reading = correlation_reading$Avg.Reading,
                      Avg.Writing = correlation_writing$Avg.Writing) 


correlationdfsorted <- reshape2::melt(correlationdf[,c("factor","Avg.Total","Avg.Math", "Avg.Reading", "Avg.Writing")],id.vars = 1)


ggplot(correlationdf, aes(reorder(factor, Avg.Total), Avg.Total)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Correlation of School Characteristics To School's Average SAT Score",
       x = "", y = "Correlation") +
  coord_flip(ylim = c(-0.75, 0.75)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
  
ggplot(correlationdfsorted, aes(reorder(factor, value), value)) +
  geom_bar(stat = "identity", aes(fill = variable), position = "dodge") +
  labs(title = "Correlation of School Characteristics To School's Average SAT Score",
       x = "", y = "Correlation") +
  coord_flip(ylim = c(-0.75, 0.75)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```



## Locations

We can see the average school results by borough, as well:

```{r, echo = F, warning = F, message = F}

#create the location df
directory = read.csv("HS_directory.csv")
locations = directory[c("dbn","school_name", "boro", "city", "zip", "Location.1")]
colnames(locations)[1] = "DBN"

# extract the the coordinates
locations$coord = gsub(".*\\((.*)\\).*", "\\1", locations$Location.1)
locations$lon <- unlist(lapply(strsplit(as.character(locations$coord), ", "), "[", 2))
locations$lat <- unlist(lapply(strsplit(as.character(locations$coord), ", "), "[", 1))
locations$lon <- as.numeric(locations$lon)
locations$lat <- as.numeric(locations$lat)

# (can map more in future using the exact coordinates extracted above)

SAT_locations = inner_join(SAT_summary_2015, locations, by = "DBN")

boros =
  SAT_locations %>%
  group_by(boro) %>%
  summarize(average = mean(Avg.Total, na.rm = TRUE), count = n()) %>%
  arrange(desc(count))

avg = ggplot(boros, mapping = aes(x=boro, y=average)) +
geom_bar(stat = "identity", col = "white", fill = "steelblue") +
  coord_flip(ylim=c(1000, 1500)) +
  labs(title = "Public Schools SAT Scores By Borough", x = "", y = "Average SAT Score for School")

count = ggplot(boros, mapping = aes(x=boro, , y=count)) +
geom_bar(stat = "identity", col = "white", fill = "steelblue") +
  coord_flip(ylim=c(0, 200)) +
  labs(title = "Number of Public Schools SAT Scores In Borough", x = "", y = "Average SAT Score for School")

grid.newpage()
grid.draw(rbind(ggplotGrob(avg), ggplotGrob(count), size = "last"))

```

Add the count!
this is not weighted average!
higher count means more public school is that borough, less resources


