Citibike Shiny (web) project:

Determine bike availability and surrounding options based on number of docked bikes

Assumptions
1. Judgement is based on prior month available pattern

Target:
1. Tourists or first time riders seeking more info on availability trends

User options
1. search for address - DONE
2. select bike availability count: - PARTIALLY DONE
  a. status of green if station docked bike count is > than user selection
  b. status of yellow if station docked bike count is == to user selection
  c. status of red if station docked bike count is < than user selection
3. select radius between 1 - 5 miles - MODIFIED: used slider feature to view surrounding bike / dock capacity
4. from and to address - produce avg ride time based on trip duration from citibike data file DONE
5. select ride time: 30 / 45 min - FUTURE ITERATION
5. select peak time view - avg bike count based on prior month data - FUTURE ITERATION

Would be nice:
1. Show available bikes and docking counts within radius of the from / to address - MODIFIED: used slider feature to view surrounding bike / dock capacity
2. When mapping station / destination, show alternative stations on the map if user goes past the alloted times (Day: 30 mins, Subscriber: 45 min) 
3. Remove gauges and add bar charts

To do: DROPPED on final version to reduce latency on web page loading
1. Download monthly citibike use data to:
  a. capture average travel time between locations
  b. avg docked bike count at stations
  c. popular stations
  d. trips within time range: 30 mins, 45 mins
2. Capture station info daily
3. Capture station status at run time
4. Build map


Citibike products:

1. Citibike:
  a. Station info / status: https://gbfs.citibikenyc.com/gbfs/gbfs.json
    i. station info: https://gbfs.citibikenyc.com/gbfs/en/station_information.json
    ii. station status: https://gbfs.citibikenyc.com/gbfs/en/station_status.json
    iii. station info / status: http://citibikenyc.com/stations/json 
  b. Monthly usage: https://s3.amazonaws.com/tripdata/index.html
  c. NYC Gov bikes: http://www.nyc.gov/html/dot/html/about/datafeeds.shtml#Bikes
2. Mobile app: https://www.citibikenyc.com/how-it-works/app
  a. provides bike availability stats per docking stations within your perimeter
3. 
  a. Leaflet Shiny how to: https://rstudio.github.io/leaflet/shiny.html
  b. https://jhonasttan.carto.com/dashboard << Carto possibilities
  c. https://mapzen.com/products/turn-by-turn/?d=0&lat=40.7259&lng=-73.9805&z=12&c=multimodal&st_lat=37.839682&st_lng=-122.485284&st=Vista%20Point&end_lat=37.80927&end_lng=-122.25981&end=Fairyland&use_bus=0.5&use_rail=0.6&use_transfers=0.4&dt=2016-10-25T08%3A00&dt_type=1 
  d. Geocoding: https://en.wikipedia.org/wiki/ISO_3166-1_alpha-2

API research and source:
1. https://cran.r-project.org/web/packages/jsonlite/vignettes/json-apis.html >>
A single public API that shows location, status and current availability for all stations in the New York City bike sharing imitative.
2. http://www.kdd.org/kdd2016/subtopic/view/rebalancing-bike-sharing-systems-a-multi-source-data-smart-optimization >> article - Rebalancing Bike Sharing Systems: A Multi-source Data Smart Optimization
3. http://toddwschneider.com/posts/a-tale-of-twenty-two-million-citi-bikes-analyzing-the-nyc-bike-share-system/ >> article - A Tale of Twenty-Two Million Citi Bike Rides: Analyzing the NYC Bike Share System
4. http://bostonography.com/hubwaymap/ - Boston Hubway (Citibike version)
5. http://www.kdd.org/kdd2016/subtopic/view/rebalancing-bike-sharing-systems-a-multi-source-data-smart-optimization << bike distribution problem
6. bike route path: http://data.beta.nyc/dataset/nyc-cycling-map
7. GPX track: https://www.r-bloggers.com/plotting-gpx-tracks-with-shiny-and-leaflet/
8. Nice shiny: http://streamstatsags.cr.usgs.gov/v3_beta/
9. Bike charts: https://spatialityblog.com/2012/05/14/citibikenyc_firstlastmile_quantified/
10. Mapzen Leaflet R: http://spatialecology.weebly.com/r-code--data/category/mapzen
11. Mapzen json request - https://mapzen.com/documentation/search/response/
12. geojson routing w/ mapzen: http://rowanwinsemius.id.au/blog/getting-started-with-routing-leaflet/
  a. routing library: https://gist.github.com/ramnathv/9998388
  b. http://stackoverflow.com/questions/28469399/not-displaying-route-using-rmaps-and-leaflet-routing-machine-plugin/33358080#33358080
  c. http://bl.ocks.org/ramnathv/9998388
  d. http://gis.stackexchange.com/questions/186748/leaflet-routing-machine-how-to-initialise-map-without-waypoints-loaded


Mapbox tutorial: https://www.mapbox.com/api-documentation/#maps
Mapbox API Playground: https://www.mapbox.com/api-playground/#/forward-geocoding/?_k=b4si62
Mapbox API Playground Directions: https://www.mapbox.com/api-documentation/#directions

Example that works:
Mapbox token: pk.eyJ1IjoiamhvbmFzdHRhbiIsImEiOiJFLTAzeVVZIn0.mwAAfKtGwv3rs3L61jz87A
https://api.mapbox.com/directions/v5/mapbox/cycling/-73.98%2C40.73%3B-73.97%2C40.75.json?access_token=pk.eyJ1IjoiamhvbmFzdHRhbiIsImEiOiJFLTAzeVVZIn0.mwAAfKtGwv3rs3L61jz87A&overview=simplified&geometries=geojson



Load libraries
```{r}
library(jsonlite)
library(dplyr)
library(cartography)
library(leaflet)
library(magrittr)
library(ggmap)
library(RCurl)
library(RJSONIO)

gGeoCode <- function(address,verbose=FALSE) {
  if(verbose) cat(address,"\n")
  u <- construct.geocode.url(address)
  doc <- getURL(u)
  x <- fromJSON(doc,simplify = FALSE)
  if(x$status=="OK") {
    lat <- x$results[[1]]$geometry$location$lat
    lng <- x$results[[1]]$geometry$location$lng
    return(c(lat, lng))
  } else {
    return(c(NA,NA))
  }
}


```

Fetch station info from REST sites:

```{r}

cb_url <- "https://feeds.citibikenyc.com/stations/stations.json"
cb_json <- fromJSON(paste(readLines(cb_url), collapse=""))
cb_stations <- cb_json$stationBeanList
cb_station_df <- data.frame(t(sapply(cb_stations,unlist)),stringsAsFactors = FALSE)

#names(cb_stations)
#nrow(cb_stations)
```

Subset of bike data from local file

```{r}

#cb_df  <- read.csv("./data/201606-citibike-tripdata.csv",stringsAsFactors = FALSE)

cb_df  <- read.csv("./data/201606-citibike-tridatasubset.csv",stringsAsFactors = FALSE) #sampled 1k rows dataset

#load("/Users/jhonasttanregalado/Documents/DataScience/bootcamp7/githubcrush/bootcamp007_project/Project2-Shiny/JhonasttanRegalado/201606-citibike-tripdata_df_duration.rds")

#which(rowSums(is.na(cb_df)) !=0 )
#length(which(rowSums(is.na(cb_df)) !=0 ))

#group_by(start.station.id, start.station.name, 
#                                                          start.station.latitude, start.station.longitude,
#                                                          end.station.id,end.station.name, end.station.latitude,
#                                                          end.station.longitude)

cb_df_avg_trip_duration <- cb_df %>% group_by(start.station.id, start.station.name, 
                                                          start.station.latitude, start.station.longitude,
                                                          end.station.id,end.station.name, end.station.latitude,
                                                          end.station.longitude) %>% 
  summarise(avg_duration_sec = mean(tripduration), 
            avg_duration_min = mean(tripduration)/60 ) %>% 
  filter(start.station.id != end.station.id) %>%  #select destinations where start station is not equal to end station.
  arrange(start.station.name) 

#saveRDS(cb_df, file = "/Users/jhonasttanregalado/Documents/DataScience/bootcamp7/githubcrush/bootcamp007_project/Project2-Shiny/JhonasttanRegalado/201606-citibike-tripdata_df.rds")
  
#saveRDS(cb_df_avg_trip_duration, file = "/Users/jhonasttanregalado/Documents/DataScience/bootcamp7/githubcrush/bootcamp007_project/Project2-Shiny/JhonasttanRegalado/201606-citibike-tripdata_df_duration.rds")

#citibike_df_avg_trip_duration_unique <- filter(citibike_df_avg_trip_duration,start.station.id != end.station.id)
```

User basic query:
1. Get from station info based on station ID
2. Get to station info based on station ID
3. Get avg duration in minutes
```{r}
s_start <- cb_stations %>% filter(id==504) %>% select(stationName,availableBikes,availableDocks, totalDocks)
s_end <- cb_stations %>% filter(id==164) %>% select(stationName,availableBikes,availableDocks, totalDocks)
avg_ride_time <- cb_df_avg_trip_duration %>% filter(start.station.id == 504, end.station.id == 164) %>% dplyr::select(avg_duration_min)
```

Map two locations - to and from

```{r}

oneRow <- cb_df_avg_trip_duration[1,c("start.station.name","start.station.latitude","start.station.longitude","end.station.name","end.station.latitude","end.station.longitude")]
tempMatrix <- oneRow %>% matrix(2,3,2)
exmapLeaflet <- as.data.frame(tempMatrix)
names(exmapLeaflet)  <- c("station.name","lat","lng")
exmapLeaflet$lat <- as.numeric(exmapLeaflet$lat)
exmapLeaflet$lng <- as.numeric(exmapLeaflet$lng)

tile_layer <- "https://api.mapbox.com/styles/v1/mapbox/streets-v10/tiles/256/{z}/{x}/{y}?access_token=pk.eyJ1IjoiamhvbmFzdHRhbiIsImEiOiJFLTAzeVVZIn0.mwAAfKtGwv3rs3L61jz87A"


conUrl <- "https://api.mapbox.com/directions/v5/mapbox/cycling/-73.98,40.73;-73.97,40.75?geometries=geojson&continue_straight=true&access_token=pk.eyJ1IjoiamhvbmFzdHRhbiIsImEiOiJFLTAzeVVZIn0.mwAAfKtGwv3rs3L61jz87A"


con <- url(conUrl)  
data.json <- fromJSON(paste(readLines(con), collapse=""))
close(con)

poly_points <- data.frame(matrix((unlist(data.json$routes[[1]]$geometry$coordinates)),length(unlist(data.json$routes[[1]]$geometry$coordinates)),2,2))
names(poly_points) <- c("lng","lat")
# rebuild poly_points with first and last row from exmapLeaflet
#poly_points <- poly_points %>% group_by(lng,lat) %>% count() %>% select(lng,lat)

map_it <- leaflet(data = exmapLeaflet) %>% 
  addTiles(urlTemplate = tile_layer) %>% 
  addPopups(exmapLeaflet, lat = exmapLeaflet$lat,lng = exmapLeaflet$lng, popup = as.character(exmapLeaflet$station.name)) %>%
  addPolylines(lng = poly_points$lng,lat = poly_points$lat,color="green", popup = "Avg Ride = 40 Mins") %>% 
  fitBounds(lng1 = exmapLeaflet$lng[1],lat1 = exmapLeaflet$lat[1], 
            lng2 = exmapLeaflet$lng[2], lat2 = exmapLeaflet$lat[2] )
map_it

#https://rstudio.github.io/leaflet/json.html


```

