---
title: "Scraping Great Thinkers"
author: "Joshua Litven"
date: "11/13/2016"
output: ioslides_presentation
---

##
<center>
*"Wisdom is nothing more profound than an ability to follow one's own advice."*

*- Sam Harris*
</center>

## Great Thinkers

- Expose us to new ideas
- Lead progress
- Change the world

## Question

What do great thinkers think about?

## Data

```{r echo=FALSE}
library(png)
library(grid)
img <- readPNG("images/pantheon.png")
grid.raster(img)
```
11,341 People spanning 2800 years across the globe.

## Data

```{r, echo=FALSE}
img <- readPNG("images/biases.png")
grid.raster(img)
```

## Data
Combine pantheon data set with quotes scraped from brainyquotes.com
```{r, echo=FALSE}
img <- readPNG("images/brainyquote.png")
grid.raster(img)
```

## Data
- Results in 4295 individuals
- Now doubly biased!
```{r, echo=FALSE}
img <- readPNG("images/missingness.png")
grid.raster(img)
```

## Insights
```{r, echo=FALSE}
img <- readPNG("images/tfidf_domain_1.png")
grid.raster(img)
```

## Insights
```{r, echo=FALSE}
img <- readPNG("images/tfidf_domain_2.png")
grid.raster(img)
```

## Insights
```{r, echo=FALSE}
img <- readPNG("images/tfidf_century_1.png")
grid.raster(img)
```

## Insights
```{r, echo=FALSE}
img <- readPNG("images/tfidf_century_2.png")
grid.raster(img)
```

## Vector Space Model
Use Vector representation of words to create a similarity matrix
```{r, eval=FALSE}
# Get most similar authors
author = "Donald Trump"

most_similar = sort(similarity_matrix[author, ], decreasing=TRUE)
most_similar = most_similar[-1] # exclude top element
most_similar[1:10]

> most_similar[1:10]
      Barack Obama        Mitt Romney  Michael Bloomberg      Rush Limbaugh          Joe Biden 
         0.3946898          0.3838377          0.3516579          0.3513637          0.3312210 
Madeleine Albright      Michael Moore      Newt Gingrich          Paul Ryan         Rick Perry 
         0.3190586          0.3116183          0.3104181          0.3067323          0.3036870
```

## Vector Space Model
Accuracy?

- Ran 1 NN Alg on author industry
- Results: 60 % accuracy (baseline is 30%)
- Conclusion: Some confidence similarity metric works

```{r, eval=FALSE}
> unique(pantheon$industry)
 [1] "GOVERNMENT"        "PHILOSOPHY"        "LANGUAGE"          "INDIVIDUAL SPORTS" "FILM AND THEATRE" 
 [6] "NATURAL SCIENCES"  "MILITARY"          "INVENTION"         "FINE ARTS"         "MATH"             
[11] "DESIGN"            "RELIGION"          "MEDICINE"          "COMPUTER SCIENCE"  "MUSIC"            
[16] "COMPANIONS"        "OUTLAWS"           "SOCIAL SCIENCES"   "BUSINESS"          "EXPLORERS"        
[21] "ACTIVISM"          "ENGINEERING"       "HISTORY"           "TEAM SPORTS"       "MEDIA PERSONALITY"
[26] "LAW"               "DANCE" 
```
## Shiny App
Demo: Ralph Waldo Emerson, William James, Donald Trump
```{r, echo=FALSE}
img <- readPNG("images/shiny.png")
grid.raster(img)
```

## Future Work
- Combine sentiment analysis with occupation, nationality and year to improve similarity metric
- Create a recommendation system

## Acknowledgements
- Fred and Nick for bouncing ideas
- Jhonasttan for project management
- CPM for telling me to go home
- Yu, A. Z., et al. (2016). Pantheon 1.0, a manually verified dataset of globally famous biographies. Scientific Data 2:150075. doi: 10.1038/sdata.2015.75

