---
title: Quotes EDA
author: "Joshua Litven"
date: "11/10/2016"
output: html_document
---

* Goal: Explore quotes of historical figures *

In this EDA we want to find the sentiments characteristic of different occupations, geographical locations and
times.

We will use term frequency inverse document frequency (TF-IDF) to distinguish between groups.

```{r setup, include=FALSE}
# knitr options
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)

# Required libraries
library(dplyr)
library(tidytext)
require(ggplot2)
library(ggstance)
library(viridis)
library(ggthemes)
```

The data used for this project: 
quotes.csv - a file scraped from brainyquotes.com containing quotes of historical figures
pantheon.csv - from the pantheon website, see https://pantheon.media.mit.edu

```{r loading data, cache=TRUE}
# Load in the quotes data set
setwd('~/Courses/nyc_data_science_academy/projects/web_scraping/data/cleaned_data')
quotes = read.csv('quotes.csv', stringsAsFactors=FALSE)
```

First we will clean the data
```{r clean, eval=FALSE }
quotes = quotes  %>% filter(author != "Lot", author !="Job", author!="Hide")
write.csv(quotes, 'quotes.csv')
```

```{r merge, cache=TRUE}
# Merge with pantheon
pantheon = read.csv('pantheon.csv', stringsAsFactors=FALSE)

quotes = merge(pantheon, quotes, by.x="name", by.y="author")

# Select columns we care about
quotes = quotes %>% 
  select(name, countryCode, continentName, birthyear, gender, occupation, industry, domain,
         body, century)
```

After loading the data, we can get the list of words using tidytext's "unnest_token" function
```{r, cache=FALSE}
# Get a list of words
words = quotes %>% unnest_tokens(word, body)
```

Now we can break up quotes into different categories to compare them.
```{r cache=TRUE}
facet_plot_tf_idf = function(group, levels_to_plot=character()){

  if(length(levels_to_plot)==0){
    levels_to_plot = unique(words[, group])
  }
  # Group words to get the counts for each group
  criteria <- lazyeval::interp(~x %in% vals, .values = list(x = as.name(group),
                                                            vals = levels_to_plot))
  words_by_group = words %>% 
    filter_(criteria) %>% 
    group_by_(group) %>% 
    count(word, sort=TRUE) %>%
    ungroup()
  
  # Combine with the total number of words in each group
  total_words = words_by_group %>% group_by_(group) %>% summarise(total = sum(n))
  words_by_group = left_join(words_by_group, total_words)
  
  # Get the TF-IDF scores of each word
  words_by_group = words_by_group %>% bind_tf_idf_(term_col="word", document_col=group, n_col="n")
  
  # Set words to factor and order by TF_IDF scores
  words_by_group =
    words_by_group %>%
    arrange(desc(tf_idf)) %>%
    mutate(word = factor(word, levels = rev(unique(word))))
  
  # Plot the top words for each group
  num_words = 10
  top_words_by_group = 
    words_by_group %>% 
    group_by_(group) %>% 
    top_n(num_words) %>% 
    ungroup
  
  ggplot(top_words_by_group, aes_string("tf_idf", "word", fill = group, alpha = "tf_idf")) +
    geom_barh(stat = "identity", show.legend = FALSE) +
    labs(title = paste("Highest TF-IDF Words by", group),
         y = NULL, x = "TF-IDF Scores") +
    facet_wrap(as.formula(paste("~", group)), ncol = 2, scales = "free") +
    theme_tufte(base_family = "Arial", base_size = 16, ticks = FALSE) +
    scale_alpha_continuous(range = c(0.6, 1)) +
    scale_x_continuous(expand=c(0,0)) +
    scale_fill_viridis(end = 0.85, discrete=TRUE) +
    theme(strip.text=element_text(hjust=0)) +
    theme(strip.text = element_text(face = "italic")) +
    theme(axis.text.x=element_blank())
}
```

View Plots of Gender, Domain, East vs. West, and millenium
```{r gender, fig.width=10, fig.height=6}
# Plot by Gender
facet_plot_tf_idf("gender")
```

```{r domain, fig.width=10, fig.height=8}
# Plot by Domain
facet_plot_tf_idf("domain")
```


```{r location, fig.width=10, fig.height=20}
# Plot by East vs West
facet_plot_tf_idf("continentName")
```

```{r time, fig.width=10, fig.height=20}
# Plot by East vs West
facet_plot_tf_idf("century")
```

