---
title: "Exploratory Data Analysis for AllState Kaggle Challenge"
output: html_document
---

# Team: Datasaurus Rex (Josh, Dina, Nick)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Let's do some basic EDA to understand our data!
```{r}
# Load in the data and summarize
library(data.table)
train = fread('../Data/train.csv')
test = fread('../Data/test.csv')
summary(train)
dim(train)
```
It looks like most categorical variables have only two levels: A and B.

All continuous variables are normalized between 0 and 1.

188318 observations
132 columns

## What does the loss function distribution look like?
```{r}
library(ggplot2)
library(gridExtra)
grid.arrange(
  ggplot(train) + geom_histogram(aes(loss), bins = 50),
  ggplot(train) + geom_histogram(aes(log(loss + 1)), bins = 50),
  ncol = 2)
```
The loss looks expontentially decreasing. 

The log(loss) looks normal.

## What do the continuous variable distributions look like?
```{r}
library(reshape2)
library(ggplot2)
library(dplyr)
cts_vars = train %>% select(contains("cont"))
cat("Number of continuous variables:", ncol(cts_vars))

melted_cts_vars <- melt(cts_vars)
ggplot(melted_cts_vars,aes(x = value)) + 
    facet_wrap(~variable,scales = "free_x") + 
    geom_histogram()
```
variable cont2 looks like discrete values. Something fishy going on!

## What are the levels in each categorical variable?
```{r}
# Output levels of categorical variables
library(dplyr)
cat_vars = train %>% select(contains("cat"))
levels_by_cat = sapply(cat_vars, levels)
levels_by_cat

# Number of categories
cat("Number of categorical variables:", ncol(cat_vars))

# Count the number of levels
cat("Number of levels Frequency Table")
plyr::count(sapply(levels_by_cat, length))
```
Most have only two levels!

## What is the correlation between the continuous variables and the loss?
```{r}
# Create correlation plot
library(corrplot)
corrs = cor(train %>% select(contains("cont"), loss), method = "pearson")
corrplot.mixed(corrs, upper = "square", order = "hclust")
```
cont9, cont6, cont10 look correlated. Nothing very correlated to loss.

## What are the t-test results for vars with two categories?
```{r}
for(var in names(levels_by_cat)){
  loss.formula = as.formula(paste("loss ~ ", var))
  if(length(levels_by_cat[[var]]) == 2){
    print(t.test(loss.formula, train))
  }else{
    print(summary(aov(loss.formula, train)))
  }
}
```
cat15, cat70 have insignificant t-tests.

cat88 has insignificant anova.



