---
title: "Spyder Project: Visualize Starbucks Store Amenities"
author: "Jhonasttan Regalado"
date: "11/7/2016"
output: html_document
---

Visualize the Starbucks locations and their amenities.
1. Scrapy to capture stores and their amenities via Starbucks Store Locator website
2. CARTO / Leaflet for visualization
3. PCA for descriptive power to determine which amenities stand out
4. Would be nice: Map layer for Shiny Citibike Analysis app

Notes:

1. Completed ability to scrape for Starbucks locations in NYC via scraper. Am getting duplicate entries, but leaving to Carto to locate based on unique geospatial info for now.
2. Created new features having amenities as yes / no using binary (1: yes, 0: no) and store names with HTML hyperlink syntax to allow user to browse Starbucks store location site

```{r}
#Load data rendered by Scrapy
library(tidyverse)
library(psych)

df_sb <- read.csv('./starbucks/StarbucksStores.txt', stringsAsFactors = FALSE)
df_sb_unique <- df_sb[!duplicated(df_sb), ]

#PCA procedure to determine which amenities should be enabled for filtering on the map
df_sb_pca <- df_sb_unique[,c(11:ncol(df_sb_unique))] #select the amenities columns
df_sb_pca <- df_sb_pca[,-ncol(df_sb_pca)] #drop the last column 'AmeOther'

head(df_sb_pca)


```
List of features:
    {"code":"LB","name":"LaBoulange"},
    {"code":"WA","name":"Oven-warmed Food"},
    {"code":"LU","name":"Lunch"},
    {"code":"CL","name":"Starbucks Reserve-Clover Brewed"},
    {"code":"WF","name":"Wireless Hotspot"},
    {"code":"CD","name":"Mobile Payment"},
    {"code":"DR","name":"Digital Rewards"},
    {"code":"GO","name":"Google Wi-Fi"},
    {"code":"FZ","name":"Fizzio Handcrafted Sodas"},
    {"code":"XO","name":"Mobile Order and Pay"},
    {"code":"RW","name":"My Starbucks Rewards"},
    {"code":"PS","name":"Playbook Store System"},
    {"code":"CS","name":"tbd - Coffee Scale"}
    {"code":"MX","name":"Music Experience"}
    {u'VS': u'Verismo'}
    {u'NB': u'Nitro Cold Brew'}
    {u'SQ': u'tbd'}
    {u'EM': u'Starbucks Evenings'}
    {u'BA': u'Bakery'}
    {u'WT': u'tbd - Walk-T'}
    {u'hrs24': u'Open 24 hours per day'}
    {u'DT': u'Drive-Through'}
    
```{r} List amenities and their sums
```{r}

df_sb_pca %>% count() #number of stores

amenMean <- map_dbl(df_sb_pca,sd)
amenMean <- sort(amenMean,decreasing = TRUE)
amenMean <- as.data.frame(amenMean, row.names = rownames(amenMean))
amenMean

amenSd <- map_dbl(df_sb_pca,sd)
amenSd <- sort(amenSd,decreasing = TRUE)
amenSd <- as.data.frame(amenSd, row.names = rownames(amenSd))
amenSd

amenSums <- map_dbl(df_sb_pca, sum)
amenSums <- sort(amenSums,decreasing = TRUE)
amenSums <- as.data.frame(amenSums, row.names = rownames(amenSums))
amenSums

amenStatsCombined <- cbind(amenSums,amenMean)
amenStatsCombined <- cbind(amenStatsCombined, amenSd)
amenStatsCombined

```
Based on frequency the: 
 - Top five most common amenities are: LB, WA, LU, DR, XO
  {"code":"LB","name":"LaBoulange"}
  {"code":"WA","name":"Oven-warmed Food"}
  {"code":"LU","name":"Lunch"}
  {"code":"DR","name":"Digital Rewards"}
  {"code":"XO","name":"Mobile Order and Pay"}

 - Bottom five least common amenities are: DT, EM, WT, hrs24, FZ
  {u'DT': u'Drive-Through'}
  {u'EM': u'Starbucks Evenings'}
  {u'WT': u'tbd - Walk-T'}
  {"code":"FZ","name":"Fizzio Handcrafted Sodas"}
  {u'hrs24': u'Open 24 hours per day'}

In the middle:

    {"code":"CL","name":"Starbucks Reserve-Clover Brewed"},
    {"code":"WF","name":"Wireless Hotspot"},
    {"code":"CD","name":"Mobile Payment"},
    {"code":"GO","name":"Google Wi-Fi"},
    {"code":"RW","name":"My Starbucks Rewards"},
    {"code":"PS","name":"Playbook Store System"},
    {"code":"CS","name":"tbd - Coffee Scale"}
    {"code":"MX","name":"Music Experience"}
    {u'VS': u'Verismo'}
    {u'NB': u'Nitro Cold Brew'}
    {u'SQ': u'tbd'}
    {u'BA': u'Bakery'}

Visual EDA: https://jhonasttan.carto.com/viz/69f64b70-a6c8-11e6-8d32-0e233c30368f/public_map 
[Visual EDA](https://jhonasttan.carto.com/viz/69f64b70-a6c8-11e6-8d32-0e233c30368f/public_map)
<<--Stop here-->>
    
PCA analysis:

```{r}
plot(df_sb_pca) # meaningless?

fa.parallel(df_sb_pca, #The data in question.
            fa = "pc", #Display the eigenvalues for PCA.
            n.iter = 100) #Number of simulated analyses to perform.
abline(h = 1)
```

Running PCA for suggested 8 PCs.
```{r}
pc_df_sb = principal(df_sb_pca, #The data in question.
                    nfactors = 8,
                    rotate = "none") #The number of PCs to extract.
pc_df_sb

```

Plotting for 8 PC does not provide a clear view of the amenities.
```{r}
factor.plot(pc_df_sb,
            labels = colnames(df_sb_pca)) #Add variable names to the plot.

plot(pc_df_sb$scores)
```

Running PCA for 3 PCs which account for 42% of the cumulative variance.
```{r}
pc_df_sb = principal(df_sb_pca, #The data in question.
                    nfactors = 3,
                    rotate = "none") #The number of PCs to extract.
pc_df_sb

```

Plotting for 3 PC does not provide a clear view of the amenities.
```{r}
factor.plot(pc_df_sb,
            labels = colnames(df_sb_pca)) #Add variable names to the plot.

plot(pc_df_sb$scores)

```

Selecting MX (Music Experience), GO (Google Wi-Fi), WF (Wireless Hotspot), RW (My Starbucks Rewards) for list of stores worth mapping.


    {"code":"CL","name":"Starbucks Reserve-Clover Brewed"},
    {"code":"WA","name":"Oven-warmed Food"},
    {"code":"WF","name":"Wireless Hotspot"},
    {"code":"CD","name":"Mobile Payment"},
    {"code":"DR","name":"Digital Rewards"},
    {"code":"LB","name":"LaBoulange"},
    {"code":"GO","name":"Google Wi-Fi"},
    {"code":"FZ","name":"Fizzio Handcrafted Sodas"},
    {"code":"XO","name":"Mobile Order and Pay"},
    {"code":"LU","name":"Lunch"},
    {"code":"RW","name":"My Starbucks Rewards"},
    {"code":"PS","name":"Playbook Store System"},
    {"code":"CS","name":"tbd - Coffee Scale"}
    {"code":"MX","name":"Music Experience"}
    {u'VS': u'Verismo'}
    {u'NB': u'Nitro Cold Brew'}
    {u'SQ': u'tbd'}
    {u'EM': u'Starbucks Evenings'}
    {u'BA': u'Bakery'}
    {u'WT': u'tbd - Walk-T'}
    {u'hrs24': u'Open 24 hours per day'}
    {u'DT': u'Drive-Through'}


Multiple Correspondence Analysis instead of PCA
http://gastonsanchez.com/how-to/2012/10/13/MCA-in-R/ 

```{r}

# load packages
require(FactoMineR)
require(ggplot2)

# load data tea
data(tea)

# select these columns
newtea = tea[, c("Tea", "How", "how", "sugar", "where", "always")]

# take a peek
head(newtea)

# number of categories per variable
cats = apply(newtea, 2, function(x) nlevels(as.factor(x)))
cats

ameCats = apply(df_sb_pca, 2, function(x) nlevels(as.factor(x)))
ameCats

# apply MCA
mca1 = MCA(newtea, graph = FALSE)
# list of results
mca1

#convert to factors
df_sb_pca[] <- lapply( df_sb_pca, factor )

# apply MCA
mcaAme = MCA(df_sb_pca, graph = FALSE)
# list of results
mcaAme
mcaAme$eig

# data frame with variable coordinates
mca1_vars_df = data.frame(mcaAme$var$coord, Variable = rep(names(ameCats), ameCats))

# data frame with observation coordinates
mca1_obs_df = data.frame(mcaAme$ind$coord)

# plot of variable categories
ggplot(data=mca1_vars_df, 
       aes(x = Dim.1, y = Dim.2, label = rownames(mca1_vars_df))) +
 geom_hline(yintercept = 0, colour = "gray70") +
 geom_vline(xintercept = 0, colour = "gray70") +
 geom_text(aes(colour=Variable)) +
 ggtitle("MCA plot of variables using R package FactoMineR")


# MCA plot of observations and categories
ggplot(data = mca1_obs_df, aes(x = Dim.1, y = Dim.2)) +
  geom_hline(yintercept = 0, colour = "gray70") +
  geom_vline(xintercept = 0, colour = "gray70") +
  geom_point(colour = "gray50", alpha = 0.7) +
  geom_density2d(colour = "gray80") +
  geom_text(data = mca1_vars_df, 
            aes(x = Dim.1, y = Dim.2, 
                label = rownames(mca1_vars_df), colour = Variable)) +
  ggtitle("MCA plot of variables using R package FactoMineR") +
  scale_colour_discrete(name = "Variable")

```

